### 2025-07-20 paper/restruct-new.tex
New

\documentclass[a4paper,english,cleveref,autoref,thm-restate]{article}

\usepackage[utf8x]{inputenc}
\usepackage{t1enc}

\usepackage{hyperref}
%\usepackage{proof}
\usepackage{fullpage}
\usepackage{amssymb}
\usepackage{stmaryrd}
%\usepackage{txfonts}
\usepackage{url}

\usepackage{xcolor}
%\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{yfonts}
\usepackage{tikz}
\usetikzlibrary{cd, arrows}

\newcommand{\Overrightarrow}[2]{\raisebox{#1}{$\ext@arrow 0359\Rightarrowfill@{\mbox{${#2}$}}{}$}}
\newcommand{\Overleftarrow}[2]{\raisebox{#1}{$\ext@arrow 0359\Leftarrowfill@{\mbox{${#2}$}}{}$}}
%\newcommand{\Overrightarrow}[1]{\raisebox{1.4ex}{$\ext@arrow 0359\Rightarrowfill@{\mbox{${#1}$}}{}$}}
\makeatother

\newcommand{\rocq}[1]{(\mbox{\texttt{#1}})}
\newcommand{\highlight}[1]{\emph{#1}}

\newcommand{\imp}{\rightarrow}
\newcommand{\Type}{\mathsf{Type}}
\newcommand{\Set}{\mathsf{Set}}
\newcommand{\Prop}{\mathsf{Prop}}
\newcommand{\refl}[1]{\mathsf{refl}\,{#1}}
\newcommand{\mkmlidtype}[3]{{#1} =^{\mathrm{id}}_{#2} {#3}}
\newcommand{\mkmleqdep}[3]{{#1} =^{\mathrm{dep}}_{#2} {#3}}
\newcommand{\hrefl}[1]{\mathsf{refl^{\mathrm{dep}}}\,{#1}}
\newcommand{\mluniv}{\mathsf{U}}
\newcommand{\Jdep}[2]{\mathsf{J}\;{#1}\;{#2}}
\newcommand{\unittype}{\ensuremath{\mathsf{unit}}}
\newcommand{\unitpoint}{\ensuremath{\ast}}

\newcommand{\nth}[1]{{#1}^{\mbox{\scriptsize th}}}
\newcommand{\defeq}{\triangleq}

\newcommand{\U}{\ensuremath{\mathsf{{HSet}}}}

\begin{document}

\section{The recursive structure of iterated parametricity}

Iterated parametricity (in indexed form) produces a family of type
families of the following form:
\begin{equation*}
  \begin{array}{llr}
    X_0 & :              & \U                                                                        \\
    X_1 & :              & X_0 \times X_0 \rightarrow  \U                                            \\
    X_2 & : \Pi a b c d. & X_1(a,b) \times X_1 (c,d) \times X_1(a,c) \times X_1 (b,d) \rightarrow \U \\
    \ldots
  \end{array}
\end{equation*}

Such family can be recursively described by the equations given on Table \ref{}.

The well-foundedness of the construction can be justified by the observation that:
\begin{enumerate}
\item Defining frame$^{n,p}$ for $p\leq n$ requires the definition of
  painting$^{n-1,k}$, thus also frame$^{n-1,k}$ by typing dependency,
  as well as restrframe$^{n,k}$, for $k < p$, where restrframe$^{n,k}$
  in turn depends in frame$^{n,k}$ for its typing. Schematically,
  using dotted arrows for dependency in the type and plain arrows for
  dependency in the definition, this gives the following dependencies:
\begin{center}
\begin{tikzcd}
frame^{n,p} \arrow[d] \arrow[ddr]\\
painting^{n-1,k<p} \arrow[d, dotted] \\
frame^{n-1,k<p} & restrframe^{n,k<p} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm]\\
\end{tikzcd}
\end{center}

In particular, the mutual dependency of frame$^{n,p}$ in the type of
restrframe$^{n,p}$ and of restrframe$^{n,p}$ in the definition of
requires to mutually defining the type of restrframe$^{n,p}$
and frame$^{n,p}$ as a function from all restrframe$^{n,k}$ for
$k < p$.

\item The same dependencies for frame$^{n-1,p}$ for $p \leq n-1$ gives:
\begin{center}
\begin{tikzcd}
frame^{n-1,p} \arrow[d] \arrow[ddr, shift right=1mm]\\
painting^{n-2,k<p} \arrow[d, dotted] \\
frame^{n-2,k<p} & restrframe^{n-1,k<p} \arrow[l, dotted] \arrow[uul, dotted, shift right=1mm]\\
\end{tikzcd}
\end{center}

Similarly, restrframe$^{n-1,p}$ and frame$^{n-1,p}$ as a function from
all restrframe$^{n-1,k}$ for $k < p$ have to be mutually defined.

\item Defining painting$^{n-1,p}$ for $p\leq n-1$, in addition to the
  type dependency in frame$^{n-1,p}$, requires the definition of
  painting$^{n-2,k}$, thus also frame$^{n-2,k}$ by typing dependency,
  as well as restrframe$^{n,k}$, this time for $k \geq p$, and the
  assumption of an inhabitant $X_{n-2}$ of $\nu$Set$^{=n-2}$
\begin{center}
\begin{tikzcd}
   & & painting^{n-1,p} \arrow[ddll] \arrow[dddl] \arrow[d, dotted]\\
   & & frame^{n-1,p} \\
painting^{n-2,k\geq p} \arrow[d, dotted] \\
frame^{n-2,k\geq p} & restrframe^{n-1,k\geq p},X_{n-2} \arrow[l, dotted]\\
\end{tikzcd}
\end{center}

\item Defining restrframe$^{n,p}$ for $p\leq n-1$, in addition to the
  type dependency in frame$^{n+1,p}$ and frame$^{n,p}$, requires
  defining restrpainting$^{n-1,k}$, thus also restrframe$^{n-1,k}$,
  frame$^{n-2,k}$ by typing dependency, as well as cohframe$^{n,k}$,
  for $k < p$, where cohframes$^{n,k}$ in turn depends in
  restrframe$^{n,k}$ for its typing:
\begin{center}
\begin{tikzcd}
frame^{n,p} \\
 \\
frame^{n-1,p} & restrframe^{n,p} \arrow[d] \arrow[ddr] \arrow[l, dotted] \arrow[uul, dotted]\\
  & restrpainting^{n-1,k<p} \arrow[d, dotted] \\
frame^{n-2,p} & restrframe^{n-1,k<p} \arrow[l, dotted] \arrow[uul, dotted] & cohframe^{n,k<p} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm]\\
\end{tikzcd}
\end{center}

In particular, the mutual dependency of restrframe$^{n,p}$ in the type
of cohframe$^{n,p}$ and of cohframe$^{n,p}$ in the definition of
restrframe$^{n,p}$ requires to mutually defining the type of
cohframe$^{n,p}$ and restrframe$^{n,p}$ as a function from all
cohframe$^{n,k}$ for $k\leq p$.

\item Defining cohframe$^{n+1,p}$ for $p\leq n-1$, in addition to the
  type dependency in restrframe$^{n+1,p}$, restrframe$^{n,p}$,
  frame$^{n+1,p}$, frame$^{n,p}$ and frame$^{n-1,p}$, requires defining
  cohpainting$^{n,k}$, thus also cohframe$^{n,k}$ by typing
  dependency, for $k < p$. By reasoning in sets, no higher-dimensional
  coherence frame is needed:
\begin{center}
\begin{tikzcd}
frame^{n+1,p} \\
\\
frame^{n,p} & restrframe^{n+1,p} \arrow[l, dotted] \arrow[uul, dotted] \\
\\
frame^{n-1,p} & restrframe^{n,p} \arrow[l, dotted] \arrow[uul, dotted] & cohframe^{n+1,p} \arrow[d] \arrow[uul, dotted] \arrow[l, dotted]\\
& & cohpainting^{n,k<p} \arrow[d, dotted] \\
frame^{n-2,p} & restrframe^{n-1,p} \arrow[l, dotted] \arrow[uul, dotted] & cohframe^{n,k<p} \arrow[uul, dotted] \arrow[l, dotted]\\
\end{tikzcd}
\end{center}
\item Defining restrpainting$^{n,p}$ for $p\leq n-1$, in addition to
  the type dependency in painting$^{n,p}$, painting$^{n-1,p}$, restrframe$^{n,p}$, frame$^{n,p}$ and
  frame$^{n-1,p}$, requires the definition of restrpainting$^{n-1,k}$,
  thus also restrframe$^{n-1,k}$ and frame$^{n-2,p}$ by typing
  dependency, as well as cohframe$^{n,k}$ for $p \geq k$ and the
  assumption of an inhabitant $X_{n-1}$ of nuSet$^{=n-1}$:
\begin{center}
\begin{tikzcd}
   & & painting^{n,p} \arrow[d, dotted] \\
   & & frame^{n,p} \\
   & & painting^{n-1,p} \arrow[d, dotted] & restrpainting^{n,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] \arrow[uul, dotted] \arrow[ul, dotted] \arrow[l, dotted] \\
   & & frame^{n-1,p} & restrframe^{n,p} \arrow[uul, dotted] \arrow[l, dotted]\\
restrpainting^{n-1,k\geq p} \arrow[d, dotted] \\
restrframe^{n-1,k\geq p} & cohframe^{n,k\geq p},X_{n-1} \arrow[l, dotted]\\
\end{tikzcd}
\end{center}

\item Defining cohpainting$^{n+1,p}$ for $p\leq n-1$, in addition to
  the type dependency in restrpainting$^{n+1,p}$,
  restrpainting$^{n,p}$, painting$^{n+1,p}$, painting$^{n,p}$,
  painting$^{n-1,p}$, restrframe$^{n+1,p}$, restrframe$^{n,p}$,
  frame$^{n+1,p}$, frame$^{n,p}$ and frame$^{n-1,p}$ requires the
  definition of cohpainting$^{n,k}$, thus also cohframe$^{n,k}$ by
  typing dependency, for also $p \geq k$. By reasoning in sets, no
  higher-dimensional coherence frame is needed:
\begin{center}
\begin{tikzcd}
   & & painting^{n+1,p} \arrow[d, dotted] \\
   & & frame^{n+1,p} \\
   & & painting^{n,p} \arrow[d, dotted] & restrpainting^{n+1,p} \arrow[d, dotted] \arrow[uul, dotted] \arrow[ul, dotted] \arrow[l, dotted] \\
   & & frame^{n,p} & restrframe^{n+1,p} \arrow[uul, dotted] \arrow[l, dotted] \\
   & & painting^{n-1,p} \arrow[d, dotted] & restrpainting^{n,p} \arrow[d, dotted] \arrow[uul, dotted] \arrow[ul, dotted] \arrow[l, dotted] & cohpainting^{n+1,p} \arrow[ddllll] \arrow[l, dotted] \arrow[uul, dotted] \arrow[d, dotted]\\
   & & frame^{n-1,p} & restrframe^{n,p} \arrow[l, dotted] & cohframe^{n+1,p} \arrow[l, dotted] \arrow[uul, dotted]\\
cohpainting^{n,k\geq p} \arrow[d, dotted] \\
cohframe^{n,k\geq p} \\
\end{tikzcd}
\end{center}
\end{enumerate}

Looking at the dependencies in the reverse order, it is thus enough:
\begin{enumerate}
\item To assume frame$^{n-2,k}$ and painting$^{n-2,k}$ for all $k\leq
  p$ in order to mutually define, for all $p\leq n$, the type of
  restrframe$^{n-1,p}$ and frame$^{n-1,p}$ as a function from all
  restrframe$^{n-1,k}$ for $k\leq p$. Using red for assumptions, this
  can be pictured as:
\begin{center}
\begin{tikzcd}
frame^{n-1,p}(\textcolor{red}{restrframe^{n-1,k<p}}) \arrow[d] \\
\textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] \\
\textcolor{red}{frame^{n-2,k<p}} \\
\end{tikzcd}
\end{center}
\item To further assume restrframe$^{n-1,k}$ for all $k\leq p$ to get
  frame$^{n-2,p}$ for all $p<n$:
\begin{center}
\begin{tikzcd}
frame^{n-1,p} \arrow[d] \arrow[ddr]\\
\textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] \\
\textcolor{red}{frame^{n-2,k<p}} & \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm]\\
\end{tikzcd}
\end{center}

\item To rely on frame$^{n-2,k}$, painting$^{n-2,k}$ and
  restrframe$^{n-1,k}$ now for all $k > p$, as well as an inhabitant
  of $X_{n-2}$ of nuSet$^{=n-2}$ to define painting$^{n-1,p}$ for all
  $p \leq n$:
\begin{center}
\begin{tikzcd}
& & & painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] \\
& & & frame^{n-1,p} \arrow[d] \arrow[ddr]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & & & \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] \\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & & \textcolor{red}{frame^{n-2,k<p}} & \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm]\\
\end{tikzcd}
\end{center}

\item To rely on frame$^{n-1,k}$ and painting$^{n-1,k}$ for all $k <
  p$ to now construct the frame$^{n,p}$ for $p \leq n-1$ as a function
  from all restrframe$^{n,k}$ for $k\leq p$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{tikzcd}
& & & frame^{n,p}(\textcolor{red}{restrframe^{n,k<p}}) \arrow[d] \\
& & & painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] \\
& & & frame^{n-1,p} \arrow[d] \\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & & & \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] \\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & & \textcolor{red}{frame^{n-2,k<p}} & \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm]\\
\end{tikzcd}

\item To additionally assume restrpainting$^{n-1,k}$ for $k < p$ to
  mutually define for $p \leq n-1$ the type of cohframe$^{n,p}$ and
  restrframe$^{n,p}$ as a function from all cohframe$^{n,k}$ for
  $k\leq p$, as well as, consequently, frame$^{n,p}$ for all $p\leq n$

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{tikzcd}
& & & frame^{n,p} \arrow[d] \arrow[ddr] \\
& & & painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] \\
& & & frame^{n-1,p} \arrow[d] \arrow[ddr] & restrframe^{n,k<p}(\textcolor{red}{cohframe^{n,k<p}}) \arrow[d] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & & & \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[d, dotted]\\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & & \textcolor{red}{frame^{n-2,k<p}} & \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm]\\
\end{tikzcd}

\item To additionally assume cohframe$^{n,k}$ for $k \leq p$ to get
  all restrframe$^{n,k}$ for all $k \leq p$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{small}
\begin{tikzcd}
& & & frame^{n,p} \arrow[d] \arrow[ddr]\\
& & & painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] \\
& & & frame^{n-1,p} \arrow[d] \arrow[ddr] & restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & & & \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted]\\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & & \textcolor{red}{frame^{n-2,k<p}} & \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] & \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm]\\
\end{tikzcd}
\end{small}

\item To rely on frame$^{n-2,k\geq p}$, painting$^{n-1,k\geq p}$ and
  restrframe$^{n,k\geq p}$ to build frame$^{n,k\geq p}$ and
  painting$^{n,k\geq p}$ for $p \leq n$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{small}
\begin{tikzcd}
& & & frame^{n,p} \arrow[d] \arrow[ddr]\\
painting^{n-1,k\geq p} \arrow[d, dotted] \arrow[dddr] \arrow[dd, bend right=70] & & & painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] \\
frame^{n-1,k\geq p} \arrow[d] \arrow[ddr] & & & frame^{n-1,p} \arrow[d] \arrow[ddr] & restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & & & \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted]\\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & & \textcolor{red}{frame^{n-2,k<p}} & \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] & \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm]\\
\end{tikzcd}
\end{small}

Note that this can be done in parallel with the steps building frame$^{n,p}$.

\item To assume restrpainting$^{n-1,k}$ and cohframe$^{n,k}$ for $k >
  p$ to build restrframe$^{n,k\geq p}$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{small}
\begin{tikzcd}
& & & frame^{n,p} \arrow[d] \arrow[ddr]\\
painting^{n-1,k\geq p} \arrow[d, dotted] \arrow[dddr] \arrow[dd, bend right=70] & & & painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] \\
frame^{n-1,k\geq p} \arrow[d] \arrow[ddr] & restrframe^{n,k\geq p} \arrow[d] \arrow[ddr] \arrow[l, dotted] & & frame^{n-1,p} \arrow[d] \arrow[ddr] & restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k\geq p}} \arrow[l, dotted] \arrow[d, dotted] & &
   \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] &
   \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted]\\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & \textcolor{red}{cohframe^{n,k\geq p},X_{n-1}} \arrow[l, dotted] &
   \textcolor{red}{frame^{n-2,k<p}} &
   \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\end{tikzcd}
\end{small}

\item To rely on painting$^{n-1,k\geq p}$ and restrframe$^{n,p\leq n-1}$ to build painting$^{n,p}$ for $p \leq n$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{small}
\begin{tikzcd}
& & & painting^{n,p} \arrow[ddlll] \arrow[dddr] \arrow[d, dotted]\\
& & & frame^{n,p} \arrow[d] \arrow[ddr]\\
painting^{n-1,k\geq p} \arrow[d, dotted] \arrow[dddr] \arrow[dd, bend right=70] & & & painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] \\
frame^{n-1,k\geq p} \arrow[d] \arrow[ddr] & restrframe^{n,k\geq p} \arrow[d] \arrow[ddr] \arrow[l, dotted] & & frame^{n-1,p} \arrow[d] \arrow[ddr] & restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k\geq p}} \arrow[l, dotted] \arrow[d, dotted] & &
   \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] &
   \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted]\\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & \textcolor{red}{cohframe^{n,k\geq p},X_{n-1}} \arrow[l, dotted] &
   \textcolor{red}{frame^{n-2,k<p}} &
   \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\end{tikzcd}
\end{small}

\item To rely on restrpainting$^{n-1,k}$ and cohframe$^{n,k}$ for $k >
  p$ to also get restrpainting$^{n,p}$ for all $p \leq n$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{small}
\begin{tikzcd}
& & & painting^{n,p} \arrow[ddlll] \arrow[dddr] \arrow[d, dotted]\\
& & & frame^{n,p} \arrow[d] \arrow[ddr]\\
painting^{n-1,k\geq p} \arrow[d, dotted] \arrow[dddr] \arrow[dd, bend right=70]  & & & painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] & restrpainting^{n,k<p} \arrow[d, dotted] \arrow[ddlll] \arrow[dddll] \arrow[l, dotted] \arrow[uul, dotted]\\
frame^{n-1,k\geq p} \arrow[d] \arrow[ddr] & restrframe^{n,k\geq p} \arrow[d] \arrow[ddr] \arrow[l, dotted] & &
  frame^{n-1,p} \arrow[d] \arrow[ddr] & restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k\geq p}} \arrow[d, dotted] & &
   \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] &
   \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted]\\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & \textcolor{red}{cohframe^{n,k\geq p},X_{n-1}} \arrow[l, dotted] &
   \textcolor{red}{frame^{n-2,k<p}} &
   \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\end{tikzcd}
\end{small}

\item To rely on frame$^{n,k}$ and painting$^{n,k}$ for all $k < p$ to
  now construct frame$^{n+1,p}$ for $p \leq n+1$ as a function from
  all restrframe$^{n+1,k}$ for $k \leq p$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{scriptsize}
\begin{tikzcd}
& & & frame^{n+1,p}(\textcolor{red}{restrframe^{n+1,k<p}}) \arrow[d]\\
& & & painting^{n,p} \arrow[ddlll] \arrow[dddr] \arrow[d, dotted]\\
& & & frame^{n,p} \arrow[d] \arrow[ddr]\\
painting^{n-1,k\geq p} \arrow[d, dotted] \arrow[dddr] \arrow[dd, bend right=70]  & & & painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] & restrpainting^{n,k<p} \arrow[d, dotted] \arrow[ddlll] \arrow[dddll] \arrow[l, dotted] \arrow[uul, dotted]\\
frame^{n-1,k\geq p} \arrow[d] \arrow[ddr] & restrframe^{n,k\geq p} \arrow[d] \arrow[ddr] \arrow[l, dotted] & &
  frame^{n-1,p} \arrow[d] \arrow[ddr] & restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k\geq p}} \arrow[d, dotted] & &
   \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] &
   \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted]\\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & \textcolor{red}{cohframe^{n,k\geq p},X_{n-1}} \arrow[l, dotted] &
   \textcolor{red}{frame^{n-2,k<p}} &
   \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\end{tikzcd}
\end{scriptsize}

\item To rely on restrframe$^{n,k}$ and restrpainting$^{n,k}$ for all
  $k < p$ to now construct restrframe$^{n+1,p}$ for $p \leq n$ as a
  function from all cohframe$^{n+1,k}$ for $k \leq p$. This also
  allows to instantiate the dependency of frame$^{n+1,p}$ on
  restrframe$^{n+1,p}$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{scriptsize}
\begin{tikzcd}
& & & frame^{n+1,p} \arrow[d] \arrow[ddr] \\
& & & painting^{n,p} \arrow[ddlll] \arrow[dddr] \arrow[d, dotted]\\
& & & frame^{n,p} \arrow[d] \arrow[ddr] & restrframe^{n+1,k<p}(\textcolor{red}{cohframe^{n+1,k<p}}) \arrow[d] \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
painting^{n-1,k\geq p} \arrow[d, dotted] \arrow[dddr] \arrow[dd, bend right=70] & & &
   painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] & restrpainting^{n,k<p} \arrow[d, dotted] \arrow[ddlll] \arrow[dddll] \arrow[l, dotted] \arrow[uul, dotted]\\
frame^{n-1,k\geq p} \arrow[d] \arrow[ddr] & restrframe^{n,k\geq p} \arrow[d] \arrow[ddr] \arrow[l, dotted] & &
  frame^{n-1,p} \arrow[d] \arrow[ddr] & restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k\geq p}} \arrow[d, dotted] & &
   \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] &
   \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted]\\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & \textcolor{red}{cohframe^{n,k\geq p},X_{n-1}} \arrow[l, dotted] &
   \textcolor{red}{frame^{n-2,k<p}} &
   \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\end{tikzcd}
\end{scriptsize}

\item To additionally assume cohpainting$^{n,k}$ for $k < p$ to get
  cohframe$^{n+1,p}$ for all $p \leq n-1$. This also allows to
  instantiate the dependency of restrframe$^{n+1,p}$ on
  cohframe$^{n+1,p}$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{scriptsize}
\begin{tikzcd}
& & & frame^{n+1,p} \arrow[d] \arrow[ddr] \\
& & & painting^{n,p} \arrow[ddlll] \arrow[dddr] \arrow[d, dotted]\\
& & & frame^{n,p} \arrow[d] \arrow[ddr] & restrframe^{n+1,k<p} \arrow[d] \arrow[ddr] \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
painting^{n-1,k\geq p} \arrow[d, dotted] \arrow[dddr] \arrow[dd, bend right=70] & & &
  painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] & restrpainting^{n,k<p} \arrow[d, dotted] \arrow[ddlll] \arrow[dddll] \arrow[l, dotted] \arrow[uul, dotted]\\
frame^{n-1,k\geq p} \arrow[d] \arrow[ddr] & restrframe^{n,k\geq p} \arrow[d] \arrow[ddr] \arrow[l, dotted] & &
 frame^{n-1,p} \arrow[d] \arrow[ddr] & restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   cohframe^{n+1,p} \arrow[d] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k\geq p}} \arrow[d, dotted] & &
   \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] &
   \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted] &
   \textcolor{red}{cohpainting^{n,k<p}} \arrow[l, dotted] \arrow[d, dotted] \\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & \textcolor{red}{cohframe^{n,k\geq p},X_{n-1}} \arrow[l, dotted] &
   \textcolor{red}{frame^{n-2,k<p}} &
   \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\end{tikzcd}
\end{scriptsize}

\item To finally assume cohpainting$^{n,k}$ for $k \geq p$
to get cohpainting$^{n+1,p}$ for $p \leq n-1$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{scriptsize}
\begin{tikzcd}
& & & frame^{n+1,p} \arrow[d] \arrow[ddr] \\
& & & painting^{n,p} \arrow[ddlll] \arrow[dddr] \arrow[d, dotted]\\
& & & frame^{n,p} \arrow[d] \arrow[ddr] & restrframe^{n+1,k<p} \arrow[d] \arrow[ddr] \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
painting^{n-1,k\geq p} \arrow[d, dotted] \arrow[dddr] \arrow[dd, bend right=70] & & &
 painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] & restrpainting^{n,k<p} \arrow[d, dotted] \arrow[ddlll] \arrow[dddll] \arrow[l, dotted] &
   cohpainting^{n+1,p} \arrow[ddlll] \arrow[d, dotted] \arrow[l, dotted] \\
frame^{n-1,k\geq p} \arrow[d] \arrow[ddr] & restrframe^{n,k\geq p} \arrow[d] \arrow[ddr] \arrow[l, dotted] & &
  frame^{n-1,p} \arrow[d] \arrow[ddr] & restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   cohframe^{n+1,p} \arrow[d] \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k\geq p}} \arrow[l, dotted] \arrow[d, dotted] & \textcolor{red}{cohpainting^{n,k\geq p}} \arrow[l, dotted] \arrow[d, dotted] &
   \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] &
   \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted] &
   \textcolor{red}{cohpainting^{n,k<p}} \arrow[l, dotted] \arrow[d, dotted] \\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & \textcolor{red}{cohframe^{n,k\geq p},X_{n-1}} \arrow[l, dotted] &
   \textcolor{red}{frame^{n-2,k<p}} &
   \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\end{tikzcd}
\end{scriptsize}
\end{enumerate}

Thus, shortly, assuming frame$^{n-2,p}$, painting$^{n-2,p}$,
restrframe$^{n-1,p}$, restrpainting$^{n-1,p}$, cohframe$^{n,p}$ and
cohpainting$^{n,p}$ for all $p \leq n-2$ allows to define
frame$^{n-1,p}$, painting$^{n-1,p}$, restrframe$^{n,p}$,
restrpainting$^{n,p}$, cohframe$^{n+1,p}$ and cohpainting$^{n+1,p}$
for all $p \leq n-1$, so the construction can be recursively
continued.

\end{document}