### 2025-08-02 paper/restruct-new.tex
Update: - attempt to define in full details the construction of frame(n)   from frame(n-1) and painting(n-1) - attempt to picture in full details the construction of painting with   dependency in extradeps - attempt to picture in full details the construction of restrframe   but avoiding dependency in extradeps

\documentclass[a4paper,english,cleveref,autoref,thm-restate]{article}

\usepackage[utf8x]{inputenc}
\usepackage{t1enc}

\usepackage{hyperref}
%\usepackage{proof}
\usepackage{fullpage}
\usepackage{amssymb}
\usepackage{stmaryrd}
%\usepackage{txfonts}
\usepackage{url}

\usepackage{xcolor}
%\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{yfonts}
\usepackage{tikz}
\usetikzlibrary{cd, arrows}

\newcommand{\Overrightarrow}[2]{\raisebox{#1}{$\ext@arrow 0359\Rightarrowfill@{\mbox{${#2}$}}{}$}}
\newcommand{\Overleftarrow}[2]{\raisebox{#1}{$\ext@arrow 0359\Leftarrowfill@{\mbox{${#2}$}}{}$}}
%\newcommand{\Overrightarrow}[1]{\raisebox{1.4ex}{$\ext@arrow 0359\Rightarrowfill@{\mbox{${#1}$}}{}$}}
\makeatother

\newcommand{\rocq}[1]{(\mbox{\texttt{#1}})}
\newcommand{\highlight}[1]{\emph{#1}}

\newcommand{\imp}{\rightarrow}
\newcommand{\Type}{\mathsf{Type}}
\newcommand{\Set}{\mathsf{Set}}
\newcommand{\Prop}{\mathsf{Prop}}
\newcommand{\refl}[1]{\mathsf{refl}\,{#1}}
\newcommand{\mkmlidtype}[3]{{#1} =^{\mathrm{id}}_{#2} {#3}}
\newcommand{\mkmleqdep}[3]{{#1} =^{\mathrm{dep}}_{#2} {#3}}
\newcommand{\hrefl}[1]{\mathsf{refl^{\mathrm{dep}}}\,{#1}}
\newcommand{\mluniv}{\mathsf{U}}
\newcommand{\Jdep}[2]{\mathsf{J}\;{#1}\;{#2}}
\newcommand{\unittype}{\ensuremath{\mathsf{unit}}}
\newcommand{\unitpoint}{\ensuremath{\ast}}

\newcommand{\nth}[1]{{#1}^{\mbox{\scriptsize th}}}
\newcommand{\defeq}{\triangleq}

\newcommand{\U}{\ensuremath{\mathsf{{HSet}}}}

\newcommand{\myframe}{\mathit{frame}}
\newcommand{\myframetype}{\mathit{frametype}}
\newcommand{\painting}{\mathit{painting}}
\newcommand{\paintingtype}{\mathit{paintingtype}}
\newcommand{\restrframe}{\mathit{restrframe}}
\newcommand{\restrframetype}{\mathit{restrframetype}}
\newcommand{\restrpainting}{\mathit{restrpainting}}
\newcommand{\cohframe}{\mathit{cohframe}}
\newcommand{\cohpainting}{\mathit{cohpainting}}

\newcommand{\udensdash}[1]{%
    \tikz[baseline=(todotted.base)]{
        \node[inner sep=1pt,outer sep=0pt] (todotted) {$#1$};
        \draw[densely dashed] (todotted.south west) -- (todotted.south east);
    }%
}%

\begin{document}

\section{The recursive structure of iterated parametricity}

Iterated parametricity (in indexed form) produces a family of type
families of the following form:
\begin{equation*}
  \begin{array}{llr}
    X_0 & :              & \U                                                                        \\
    X_1 & :              & X_0 \times X_0 \rightarrow  \U                                            \\
    X_2 & : \Pi a b c d. & X_1(a,b) \times X_1 (c,d) \times X_1(a,c) \times X_1 (b,d) \rightarrow \U \\
    \ldots
  \end{array}
\end{equation*}

Such family can be recursively described by the equations given on Table \ref{}.

The well-foundedness of the construction can be justified by the observation that:
\begin{enumerate}
\item Defining $\myframe^{n,p}$ for $p \leq n$ requires the definition
  of $\painting^{n-1,p-1}$, thus also $\myframe^{n-1,p-1}$ by typing
  dependency, as well as $\restrframe^{n,p-1}$, where
  $\restrframe^{n,p-1}$ in turn depends in $\myframe^{n,p-1}$ for its
  typing.

  Moreover, $\myframe^{n,p}$ depends on $\myframe^{n,p-1}$, so these
  dependencies are between the blocks of $\myframe^{n,k}$ for all
  $k\leq p$, as well as $\myframe^{n-1,k}$, $\painting^{n-1,k}$ and
  $\restrframe^{n,k}$ for all $k\leq p-1$, which we respectively
  abbreviate as $\myframe^{n,[0,...,p]}$,
  $\myframe^{n-1,[0,...,p-1]}$, $\painting^{n-1,[0,...,p-1]}$ and
  $\restrframe^{n,[0,...,p-1]}$.

  Also, $\restrframe^{n,p-1}$ depends on $n$.

  Schematically, using dotted arrows for dependencies in
  the type and plain arrows for dependencies in the definition, this
  gives the following dependencies for $p\leq n$ fixed:
\begin{center}
\begin{tikzcd}
\myframe^{n,[0,...,p]} \arrow[d] \arrow[ddr]\\
\painting^{n-1,[0,...,p-1]} \arrow[d, dotted] \\
\myframe^{n-1,[0,...,p-1]} &
   \restrframe^{n,[0,...,p-1]} \arrow[l, dotted] \arrow[uul, dotted, "\mbox{$[0,...,p-1]$}"', near start, shift right=2mm, hook']\\
\end{tikzcd}
\end{center}

In particular, the mutual dependency on one side of
$\myframe^{n,[0,...,p-1]}$ within the type of
$\restrframe^{n,[0,...,p-1]}$ and on the other side of
$\restrframe^{n,[0,...,p-1]}$ within the definition of
$\myframe^{n,[0,...,p]}$ requires to mutually define the type of
$\restrframe^{n,[0,...,p-1]}$ together with $\myframe^{n,p}$ (or
$\myframe^{n,[0,...,p]}$) as a function from
$\restrframe^{n,[0,...,p-1]}$.

To talk about sequences of $\myframe$, $\painting$ or $\restrframe$,
we need to specify their types. The type of $\restrframe$ will be
defined mutually with $\myframe$ of the same level, but the type of
$\myframe$, $\painting$ can be defined in advance, the second one
being however dependent on the first one. In general, components of
such sequences will have a type dependent on the previous components
of the sequence, as it is the case for $\restrframe$. So, the type of
a sequence of length $p$ will be expressed as a $n$-iterated
$\Sigma$-type. In the case of $\myframe$ and $\painting$, there will
be however no dependency, so an ordinary product can be used. If
$\myframe^{n,[0,...,p]}$ is such a sequence of length $p+1$, we write
$\myframe^{n,[0,...,p-1]}$ for its immediate subsequence of length $p$
and $\myframe^{n,p}$ for its last component. We can now define the
types of sequences of $\myframe$ and $\painting$ as follows:

$$
\begin{array}{llcl}
\myframetype^{n,[0,...,p]} & & : & Type_{l+1} \\
\myframetype^{n,[]} & & \defeq & unit \\
\myframetype^{n,[0,...,p]} & & \defeq & \myframetype^{n,[0,...,p-1]} \times Type_l\\
\\
\paintingtype^{n,[0,...,p]} & (\myframe^{n,[0,...,p]}:\myframetype^{n,[0,...,p]}) & : & Type_{l+1} \\
\paintingtype^{n,[]} & \star & \defeq & unit \\
\paintingtype^{n,[0,...,p]} & (\myframe^{n,[0,...,p]}) & \defeq & \paintingtype^{n,[0,...,p-1]} \times (\myframe^{n,p} \rightarrow Type_l)\\
\end{array}
$$

Formally, this can be described as:
$$
\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!
\begin{array}{llcl}
\myframe^{n,[0,...,p]} &
  \multicolumn{3}{l}{
  \left(\begin{array}{l}
     \myframe^{n-1,[0,...,p-1]}:\myframetype^{n-1,[0,...,p-1]}\\
     \painting^{n-1,[0,...,p-1]}:\paintingtype^{n-1,[0,...,p-1]}(\myframe^{n-1,[0,...,p-1]})\\
     \restrframe^{n,[0,...,p-1]}:\restrframetype^{n,[0,...,p-1]}\\
  \end{array}\right) : \myframetype^{n-1,[0,...,p]}
  }\\
\myframe^{n,[0]} &
   \left(\begin{array}{l}
      \star\\
      \star\\
      \star\\
   \end{array}\right) & \defeq & (\star,unit)\\
\myframe^{n,[0,...,p+1]} &
   \left(\begin{array}{l}
      \myframe^{n-1,[0,...,p]}\\
      \painting^{n-1,[0,...,p]}\\
     \restrframe^{n,[0,...,p]}\\
   \end{array}\right) & \defeq &
   \left(\begin{array}{l}
     \myframe^{n,[0,...,p]}
             \left(\begin{array}{l}
               \myframe^{n-1,[0,...,p-1]}\\
               \painting^{n-1,[0,...,p-1]}\\
               \restrframe^{n,[0,...,p-1]}\\
             \end{array}\right)     ,\\
     \Sigma d:\myframe^{n,[0,...,p]}
             \left(\begin{array}{l}
               \myframe^{n-1,[0,...,p-1]}\\
               \painting^{n-1,[0,...,p-1]}\\
               \restrframe^{n,[0,...,p-1]}\\
             \end{array}\right).\,\\
     \quad \Pi\omega.\,\painting^{n-1,p}(\restrframe^{n,p}(d))\\
   \end{array}\right)\\
\restrframetype^{n,[0,...,p-1]} &
   \multicolumn{3}{l}{
   \left(\begin{array}{l}
      \myframe^{n-1,[0,...,p-1]}:\myframetype^{n-1,[0,...,p-1]}\\
      \painting^{n-1,[0,...,p-1]}:\paintingtype^{n-1,[0,...,p-1]}(\myframe^{n-1,[0,...,p-1]})\\
   \end{array}\right) : Type_{l}
   }\\
\restrframetype^{n,[]} &
   \left(\begin{array}{l}
      \star\\
      \star\\
   \end{array}\right) & \defeq &
   unit\\
\restrframetype^{n,[0,...,p]} &
   \left(\begin{array}{l}
      \myframe^{n-1,[0,...,p]}\\
      \painting^{n-1,[0,...,p]}\\
   \end{array}\right) & \defeq &
   \begin{array}{l}
     \Sigma \restrframe^{n,[0,...,p-1]}\\
     \quad:\restrframetype^{n,[0,...,p-1]}
                \left(\begin{array}{l}
                  \myframe^{n-1,[0,...,p-1]}\\
                  \painting^{n-1,[0,...,p-1]}\\
                \end{array}\right)
                .\,\\
     \Pi q\leq n.\,\Pi \omega.\,\myframe^{n,[0,...,p]}
       \left(\begin{array}{l}
         \myframe^{n-1,[0,...,p-1]}\\
         \painting^{n-1,[0,...,p-1]}\\
         \restrframe^{n,[0,...,p-1]}\\
       \end{array}\right)\\
   \end{array}\\
\end{array}
$$

TODO: discuss $p\leq q \leq n$ vs $0 \leq q \leq n-p$.

TODO: remark that we abuse terminology for frame, since, informally, a
frame is an inhabitant of the type frame, while formally, it is the
type of (an informal) frame which we call frame. (See same abuse of
terminology when talking about a list, and using the term list to
refer to the type of a list.)

TODO: when to talk about the ``prefix'' $X_0, ..., X_{n-1}$?

%% \item By induction on $p$, we deduce this for all $p\leq n$:
%% \begin{center}
%% \begin{tikzcd}
%% frame^{n,[0,...,p...n]} \arrow[d] \arrow[ddr]\\
%% painting^{n-1,[0,...,p-1...n-1]} \arrow[d, dotted] \\
%% frame^{n-1,[0,...,p-1...n-1]} &
%%    \restrframe^{n,[0,...,p-1...n-1]} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm]\\
%% \end{tikzcd}
%% \end{center}

%% \item The same dependencies for $\myframe^{n-1,p}$ for $p \leq n-1$ gives:
%% \begin{center}
%% \begin{tikzcd}
%% frame^{n-1,p} \arrow[d] \arrow[ddr, shift right=1mm]\\
%% painting^{n-2,k<p} \arrow[d, dotted] \\
%% frame^{n-2,k<p} & \restrframe^{n-1,k<p} \arrow[l, dotted] \arrow[uul, dotted, shift right=1mm]\\
%% \end{tikzcd}
%% \end{center}

%% Similarly, $\restrframe^{n-1,p}$ and $\myframe^{n-1,p}$ as a function from
%% all $\restrframe^{n-1,k}$ for $k < p$ have to be mutually defined.

\item Defining $\painting^{n,p}$, in addition to the type dependency
  in $\myframe^{n,p}$, requires the definition of
  $\painting^{n-1,p-1}$, thus also $\myframe^{n-1,p-1}$ by typing
  dependency, as well as $\restrframe^{n,p-1}$. Since
  $\painting^{n,p}$ also depends on $\painting^{n,p+1}$ when $p<n$,
  these dependencies are actually between $\painting^{n,[p,...,n]}$,
  $\painting^{n-1,[p,...,n-1]}$, $\myframe^{n,[p,...,n]}$,
  $\myframe^{n-1,[p,...,n-1]}$ and $\restrframe^{n,[p,...,n]}$. Also,
  for the case $p=n$, this also requires the assumption of an
  inhabitant $X_{n}$ of $\nu$Set$^{=n}$. Finally,
  $\myframe^{n,[p,...,n]}$, $\myframe^{n-1,[p,...,n-1]}$ and
  $\restrframe^{n,[p,...,n-1]}$ respectively depend on the earlier
  values $\myframe^{n,[0,...,p-1]}$, $\myframe^{n-1,[0,...,p-1]}$ and
  $\restrframe^{n,[0,...,p-1]}$.  Below, we use dashed underline and
  dashed arrows for dependencies in subconstructions, here in the
  construction of $\myframe$ at level $n-1$ and $\restrframetype$ at
  level $n$:
\begin{center}
\begin{tikzcd}
& & \painting^{n,[p,...,n]} \arrow[dd, bend right=70] \arrow[dddr] \arrow[d, dotted]\\
\udensdash{\myframe}^{n,[0,...,p-1]} \arrow[d, dashed] \arrow[ddr, dashed] & &
  \udensdash{\myframe}^{n,[p,...,n]} \arrow[ll, dotted] \arrow[d, dashed] \arrow[ddr, dashed, shift right=2mm] \\
\painting^{n-1,[0,...,p-1]} \arrow[d, dotted] & &
  \painting^{n-1,[p,...,n-1]} \arrow[ll, dotted] \arrow[d, dotted] \\
\myframe^{n-1,[0,...,p-1]} &
  \udensdash{\restrframe}^{n,[0,...,p-1]} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm, hook'] &
  \myframe^{n-1,[p,...,n-1]} \arrow[ll, dotted, bend left=15] &
  \udensdash{\restrframe}^{n,[p,...,n-1]},X_n \arrow[ll, dotted, bend left=15] \arrow[l, dotted] \arrow[uul, dotted, hook']\\
\end{tikzcd}
\end{center}

\item Defining $\restrframe^{n,p}$ for $p\leq n-1$, in addition to the
  mutual type dependency in $\myframe^{n,p}$ and the dependency in
  $\myframe^{n-1,p}$, requires $\restrpainting^{n-1,p-1}$, thus also
  $\restrframe^{n-1,p-1}$ and $\myframe^{n-1,p-1}$ mutually, as well
  as $\myframe^{n-2,p-2}$, $painting^{n-2,p-1}$ and
  $\painting^{n-2,p-1}$ by typing dependency, as well as
  $\cohframe^{n,p-1}$. Since $\restrframe^{n,p}$ also depends on
  $\restrframe^{n,p-1}$, these dependencies are between the blocks
  $\myframe^{n,[0..p]}$, $\myframe^{n-1,[0..p]}$,
  $\myframe^{n-2,[0..p]}$, $\painting^{n-1,[0..p-1]}$ and
  $\restrframe^{n,[0..p-1]}$, where cohframes$^{n,k}$ in turn depends
  in $\restrframe^{n,k}$ for its typing. Below, we use again dashed
  underline and dashed arrows for dependencies in subconstructions,
  here in the construction of $\myframe$ at level $n-1$ and $n-2$ and $\restrframetype$ at level $n$ and $n-1$ :

\begin{center}
\begin{tikzcd}
\udensdash{\myframe}^{n,[0..p]} \arrow[d, dashed] \arrow[ddr, dashed] \\
\painting^{n-1,[0..p-1]} \arrow[d, dotted]\\
\udensdash{\myframe}^{n-1,[0..p-1]} \arrow[d, dashed] \arrow[ddr, dashed] &
  \restrframe^{n,[0..p-1]} \arrow[d] \arrow[ddr] \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm, hook']\\
\painting^{n-2,[0..p-2]} \arrow[d, dotted] &
  \restrpainting^{n-1,[0..p-2]} \arrow[d, dotted] \arrow[l, dotted] \arrow[uul, dotted, hook']\\
\myframe^{n-2,[0..p-2]} &
  \restrframe^{n-1,[0..p-2]} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm, hook'] &
  \cohframe^{n,[0..p-2]} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm, hook']\\
\end{tikzcd}
\end{center}

In particular, the mutual dependency of $\restrframe^{n,[0,...,p-1]}$
in the type of $\cohframe^{n,[0,...,p-2]}$ and of
$\cohframe^{n,[0,..,p-2]}$ in the definition of
$\restrframe^{n,[0,...,p-1]}$ requires to mutually define the type of
$\cohframe^{n,[0,...,p-2]}$ and $\restrframe^{n,p}$
($\restrframe^{n,[0,...,p-1]}$) as a function from
  $\cohframe^{n,[0,...,p-2]}$.

\item Defining $\restrpainting^{n,p}$ for $p\leq n-1$, in addition to
  the type dependency in $\painting^{n,p}$, $\painting^{n-1,p}$, $\restrframe^{n,p}$, $\myframe^{n,p}$ and
  $\myframe^{n-1,p}$, requires the definition of $\restrpainting^{n-1,k}$,
  thus also $\restrframe^{n-1,k}$ and $\myframe^{n-2,p}$ by typing
  dependency, as well as $\cohframe^{n,k}$ for $p \geq k$ and the
  assumption of an inhabitant $X_{n-1}$ of nuSet$^{=n-1}$:
\begin{center}
\begin{tikzcd}
   & & painting^{n,p} \arrow[d, dotted] \\
   & & frame^{n,p} \\
   & & painting^{n-1,p} \arrow[d, dotted] & \restrpainting^{n,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] \arrow[uul, dotted] \arrow[ul, dotted] \arrow[l, dotted] \\
   & & frame^{n-1,p} & \restrframe^{n,p} \arrow[uul, dotted] \arrow[l, dotted]\\
restrpainting^{n-1,[p,...,n]} \arrow[d, dotted] \\
restrframe^{n-1,[p,...,n]} & \cohframe^{n,[p,...,n]},X_{n-1} \arrow[l, dotted]\\
\end{tikzcd}
\end{center}

\item Defining $\cohframe^{n+1,p}$ for $p\leq n-1$, in addition to the
  type dependency in $\restrframe^{n+1,p}$, $\restrframe^{n,p}$,
  $\myframe^{n+1,p}$, $\myframe^{n,p}$ and $\myframe^{n-1,p}$, requires defining
  $\cohpainting^{n,k}$, thus also $\cohframe^{n,k}$, $\restrpainting^{n,k}$, $\restrpainting^{n-1,k}$, $\painting^{n,k}$, $\painting^{n-1,k}$ and $\painting^{n-2,k}$ by typing
  dependency, for $k < p$. By reasoning in sets, no higher-dimensional
  coherence frame is needed:
\begin{center}
\begin{tikzcd}
\myframe^{n+1,p} \\
\painting^{n,p} \arrow[d, dotted] \\
\myframe^{n,p} & \restrframe^{n+1,p} \arrow[l, dotted] \arrow[uul, dotted, hook'] \\
\painting^{n-1,p} \arrow[d, dotted] &
  \restrpainting^{n-1,p} \arrow[d, dotted] \arrow[l, dotted] \arrow[uul, dotted, hook'] \\
\myframe^{n-1,p} &
  \restrframe^{n,p} \arrow[l, dotted] \arrow[uul, dotted, hook'] &
  \cohframe^{n+1,p} \arrow[d] \arrow[l, dotted] \arrow[uul, dotted, hook'] \\
\painting^{n-2,p} \arrow[d, dotted] &
  \restrpainting^{n-1,p} \arrow[d, dotted] \arrow[l, dotted] \arrow[uul, dotted, hook'] &
  \cohpainting^{n,k<p} \arrow[d, dotted] \arrow[l, dotted] \arrow[uul, dotted, hook']\\
\myframe^{n-2,p} &
  \restrframe^{n-1,p} \arrow[l, dotted] \arrow[uul, dotted, hook'] &
  \cohframe^{n,k<p} \arrow[l, dotted] \arrow[uul, dotted, hook']\\
\end{tikzcd}
\end{center}

\item Defining $\cohpainting^{n+1,p}$ for $p\leq n-1$, in addition to
  the type dependency in $\restrpainting^{n+1,p}$,
  $\restrpainting^{n,p}$, $\painting^{n+1,p}$, $\painting^{n,p}$,
  $\painting^{n-1,p}$, $\restrframe^{n+1,p}$, $\restrframe^{n,p}$,
  $\myframe^{n+1,p}$, $\myframe^{n,p}$ and $\myframe^{n-1,p}$ requires the
  definition of $\cohpainting^{n,k}$, thus also $\cohframe^{n,k}$ by
  typing dependency, for also $p \geq k$. By reasoning in sets, no
  higher-dimensional coherence frame is needed:
\begin{center}
\begin{tikzcd}
   & & painting^{n+1,p} \arrow[d, dotted] \\
   & & frame^{n+1,p} \\
   & & painting^{n,p} \arrow[d, dotted] & \restrpainting^{n+1,p} \arrow[d, dotted] \arrow[uul, dotted] \arrow[ul, dotted] \arrow[l, dotted] \\
   & & frame^{n,p} & \restrframe^{n+1,p} \arrow[uul, dotted] \arrow[l, dotted] \\
   & & painting^{n-1,p} \arrow[d, dotted] & \restrpainting^{n,p} \arrow[d, dotted] \arrow[uul, dotted] \arrow[ul, dotted] \arrow[l, dotted] & \cohpainting^{n+1,p} \arrow[ddllll] \arrow[l, dotted] \arrow[uul, dotted] \arrow[d, dotted]\\
   & & frame^{n-1,p} & \restrframe^{n,p} \arrow[l, dotted] & \cohframe^{n+1,p} \arrow[l, dotted] \arrow[uul, dotted]\\
cohpainting^{n,k\geq p} \arrow[d, dotted] \\
cohframe^{n,k\geq p} \\
\end{tikzcd}
\end{center}
\end{enumerate}

Looking at the dependencies in the reverse order, it is thus enough:
\begin{enumerate}
\item To assume $\myframe^{n-2,k}$ and $\painting^{n-2,k}$ for all $k< p$
  in order to mutually define, for all $p\leq n$, the type of
  $\restrframe^{n-1,p}$ and $\myframe^{n-1,p}$ as a function from all
  $\restrframe^{n-1,k}$ for $k< p$. Using red for assumptions, this
  can be pictured as:
\begin{center}
\begin{tikzcd}
\myframe^{n-1,p}(\textcolor{red}{restrframe^{n-1,k<p}}) \arrow[d] \\
\textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] \\
\textcolor{red}{frame^{n-2,k<p}} \\
\end{tikzcd}
\end{center}

\item To further assume $\restrframe^{n-1,k}$ for all $k<p$ to get
  $\myframe^{n-1,p}$ for all $p\leq n-1$:
\begin{center}
\begin{tikzcd}
\myframe^{n-1,p} \arrow[d] \arrow[ddr]\\
\textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] \\
\textcolor{red}{frame^{n-2,k<p}} & \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm]\\
\end{tikzcd}
\end{center}

\item To rely on $\myframe^{n-2,k}$, $\painting^{n-2,k}$ and
  $\restrframe^{n-1,k}$, jointly with $\myframe^{n-1,k}$, and now for all
  $k \geq p$, as well as an inhabitant $X_{n-2}$ of $\nu$Set$^{=n-2}$
  to define $\painting^{n-1,p}$ for all $p \leq n-1$:
\begin{center}
\begin{tikzcd}
& & & painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] \\
\myframe^{n-1,k\geq p} \arrow[d] \arrow[ddr] & & & frame^{n-1,p} \arrow[d] \arrow[ddr]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & & & \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] \\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & & \textcolor{red}{frame^{n-2,k<p}} & \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm]\\
\end{tikzcd}
\end{center}

\item To rely on $\myframe^{n-1,k}$ and $\painting^{n-1,k}$ for all $k <
  p$ to now construct the $\myframe^{n,p}$ for $p \leq n-1$ as a function
  from all $\restrframe^{n,k}$ for $k\leq p$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{tikzcd}
& & & frame^{n,p}(\textcolor{red}{restrframe^{n,k<p}}) \arrow[d] \\
& & & painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] \\
& & & frame^{n-1,p} \arrow[d] \\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & & & \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] \\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & & \textcolor{red}{frame^{n-2,k<p}} & \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm]\\
\end{tikzcd}

\item To additionally assume $\restrpainting^{n-1,k}$ for $k < p$ to
  mutually define for $p \leq n-1$ the type of $\cohframe^{n,p}$ and
  $\restrframe^{n,p}$ as a function from all $\cohframe^{n,k}$ for
  $k\leq p$, as well as, consequently, $\myframe^{n,p}$ for all $p\leq n$

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{tikzcd}
& & & frame^{n,p} \arrow[d] \arrow[ddr] \\
& & & painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] \\
& & & frame^{n-1,p} \arrow[d] \arrow[ddr] & \restrframe^{n,k<p}(\textcolor{red}{cohframe^{n,k<p}}) \arrow[d] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & & & \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[d, dotted]\\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & & \textcolor{red}{frame^{n-2,k<p}} & \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm]\\
\end{tikzcd}

\item To additionally assume $\cohframe^{n,k}$ for $k \leq p$ to get
  all $\restrframe^{n,k}$ for all $k \leq p$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{small}
\begin{tikzcd}
& & & frame^{n,p} \arrow[d] \arrow[ddr]\\
& & & painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] \\
& & & frame^{n-1,p} \arrow[d] \arrow[ddr] & \restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & & & \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted]\\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & & \textcolor{red}{frame^{n-2,k<p}} & \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] & \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm]\\
\end{tikzcd}
\end{small}

\item To rely on $\myframe^{n-2,k\geq p}$, $\painting^{n-2,k\geq p}$ and
  $\restrframe^{n-1,k\geq p}$ to build $\myframe^{n-1,k\geq p}$ and
  $\painting^{n-1,k\geq p}$ for $p \leq n$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{small}
\begin{tikzcd}
& & & frame^{n,p} \arrow[d] \arrow[ddr]\\
\painting^{n-1,k\geq p} \arrow[d, dotted] \arrow[dddr] \arrow[dd, bend right=70] & & & painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] \\
\myframe^{n-1,k\geq p} \arrow[d] \arrow[ddr] & & & frame^{n-1,p} \arrow[d] \arrow[ddr] & \restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & & & \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted]\\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & & \textcolor{red}{frame^{n-2,k<p}} & \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] & \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm]\\
\end{tikzcd}
\end{small}

Note that this can be done in parallel with the steps building $\myframe^{n,p}$.

\item To assume $\restrpainting^{n-1,k}$ and $\cohframe^{n,k}$ for $k >
  p$ to build $\restrframe^{n,k\geq p}$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{small}
\begin{tikzcd}
& & & frame^{n,p} \arrow[d] \arrow[ddr]\\
\painting^{n-1,k\geq p} \arrow[d, dotted] \arrow[dddr] \arrow[dd, bend right=70] & & & painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] \\
\myframe^{n-1,k\geq p} \arrow[d] \arrow[ddr] & \restrframe^{n,k\geq p} \arrow[d] \arrow[ddr] \arrow[l, dotted] & & frame^{n-1,p} \arrow[d] \arrow[ddr] & \restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k\geq p}} \arrow[l, dotted] \arrow[d, dotted] & &
   \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] &
   \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted]\\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & \textcolor{red}{cohframe^{n,k\geq p},X_{n-1}} \arrow[l, dotted] &
   \textcolor{red}{frame^{n-2,k<p}} &
   \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\end{tikzcd}
\end{small}

\item To rely on $\painting^{n-1,k\geq p}$ and $\restrframe^{n,p\leq n-1}$ to build $\painting^{n,p}$ for $p \leq n$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{small}
\begin{tikzcd}
& & & painting^{n,p} \arrow[ddlll] \arrow[dddr] \arrow[d, dotted]\\
& & & frame^{n,p} \arrow[d] \arrow[ddr]\\
\painting^{n-1,k\geq p} \arrow[d, dotted] \arrow[dddr] \arrow[dd, bend right=70] & & & painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] \\
\myframe^{n-1,k\geq p} \arrow[d] \arrow[ddr] & \restrframe^{n,k\geq p} \arrow[d] \arrow[ddr] \arrow[l, dotted] & & frame^{n-1,p} \arrow[d] \arrow[ddr] & \restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k\geq p}} \arrow[l, dotted] \arrow[d, dotted] & &
   \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] &
   \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted]\\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & \textcolor{red}{cohframe^{n,k\geq p},X_{n-1}} \arrow[l, dotted] &
   \textcolor{red}{frame^{n-2,k<p}} &
   \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\end{tikzcd}
\end{small}

\item To rely on $\restrpainting^{n-1,k}$ and $\cohframe^{n,k}$ for $k >
  p$ to also get $\restrpainting^{n,p}$ for all $p \leq n-1$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{small}
\begin{tikzcd}
& & & painting^{n,p} \arrow[ddlll] \arrow[dddr] \arrow[d, dotted]\\
& & & frame^{n,p} \arrow[d] \arrow[ddr]\\
\painting^{n-1,k\geq p} \arrow[d, dotted] \arrow[dddr] \arrow[dd, bend right=70]  & & & painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] & \restrpainting^{n,k<p} \arrow[d, dotted] \arrow[ddlll] \arrow[dddll] \arrow[l, dotted] \arrow[uul, dotted]\\
\myframe^{n-1,k\geq p} \arrow[d] \arrow[ddr] & \restrframe^{n,k\geq p} \arrow[d] \arrow[ddr] \arrow[l, dotted] & &
  frame^{n-1,p} \arrow[d] \arrow[ddr] & \restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k\geq p}} \arrow[d, dotted] & &
   \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] &
   \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted]\\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & \textcolor{red}{cohframe^{n,k\geq p},X_{n-1}} \arrow[l, dotted] &
   \textcolor{red}{frame^{n-2,k<p}} &
   \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\end{tikzcd}
\end{small}

\item To rely on $\myframe^{n,k}$ and $\painting^{n,k}$ for all $k < p$ to
  now construct $\myframe^{n+1,p}$ for $p \leq n+1$ as a function from
  all $\restrframe^{n+1,k}$ for $k \leq p$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{scriptsize}
\begin{tikzcd}
& & & frame^{n+1,p}(\textcolor{red}{restrframe^{n+1,k<p}}) \arrow[d]\\
& & & painting^{n,p} \arrow[ddlll] \arrow[dddr] \arrow[d, dotted]\\
& & & frame^{n,p} \arrow[d] \arrow[ddr]\\
\painting^{n-1,k\geq p} \arrow[d, dotted] \arrow[dddr] \arrow[dd, bend right=70]  & & & painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] & \restrpainting^{n,k<p} \arrow[d, dotted] \arrow[ddlll] \arrow[dddll] \arrow[l, dotted] \arrow[uul, dotted]\\
\myframe^{n-1,k\geq p} \arrow[d] \arrow[ddr] & \restrframe^{n,k\geq p} \arrow[d] \arrow[ddr] \arrow[l, dotted] & &
  frame^{n-1,p} \arrow[d] \arrow[ddr] & \restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k\geq p}} \arrow[d, dotted] & &
   \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] &
   \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted]\\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & \textcolor{red}{cohframe^{n,k\geq p},X_{n-1}} \arrow[l, dotted] &
   \textcolor{red}{frame^{n-2,k<p}} &
   \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\end{tikzcd}
\end{scriptsize}

\item To rely on $\restrframe^{n,k}$ and $\restrpainting^{n,k}$ for all
  $k < p$ to now construct $\restrframe^{n+1,p}$ for $p \leq n$ as a
  function from all $\cohframe^{n+1,k}$ for $k \leq p$. This also
  allows to instantiate the dependency of $\myframe^{n+1,p}$ on
  $\restrframe^{n+1,p}$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{scriptsize}
\begin{tikzcd}
& & & frame^{n+1,p} \arrow[d] \arrow[ddr] \\
& & & painting^{n,p} \arrow[ddlll] \arrow[dddr] \arrow[d, dotted]\\
& & & frame^{n,p} \arrow[d] \arrow[ddr] & \restrframe^{n+1,k<p}(\textcolor{red}{cohframe^{n+1,k<p}}) \arrow[d] \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\painting^{n-1,k\geq p} \arrow[d, dotted] \arrow[dddr] \arrow[dd, bend right=70] & & &
   painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] & \restrpainting^{n,k<p} \arrow[d, dotted] \arrow[ddlll] \arrow[dddll] \arrow[l, dotted] \arrow[uul, dotted]\\
\myframe^{n-1,k\geq p} \arrow[d] \arrow[ddr] & \restrframe^{n,k\geq p} \arrow[d] \arrow[ddr] \arrow[l, dotted] & &
  frame^{n-1,p} \arrow[d] \arrow[ddr] & \restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k\geq p}} \arrow[d, dotted] & &
   \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] &
   \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted]\\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & \textcolor{red}{cohframe^{n,k\geq p},X_{n-1}} \arrow[l, dotted] &
   \textcolor{red}{frame^{n-2,k<p}} &
   \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\end{tikzcd}
\end{scriptsize}

\item To additionally assume $\cohpainting^{n,k}$ for $k < p$ to get
  $\cohframe^{n+1,p}$ for all $p \leq n-1$. This also allows to
  instantiate the dependency of $\restrframe^{n+1,p}$ on
  $\cohframe^{n+1,p}$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{scriptsize}
\begin{tikzcd}
& & & frame^{n+1,p} \arrow[d] \arrow[ddr] \\
& & & painting^{n,p} \arrow[ddlll] \arrow[dddr] \arrow[d, dotted]\\
& & & frame^{n,p} \arrow[d] \arrow[ddr] & \restrframe^{n+1,k<p} \arrow[d] \arrow[ddr] \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\painting^{n-1,k\geq p} \arrow[d, dotted] \arrow[dddr] \arrow[dd, bend right=70] & & &
  painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] & \restrpainting^{n,k<p} \arrow[d, dotted] \arrow[ddlll] \arrow[dddll] \arrow[l, dotted] \arrow[uul, dotted]\\
\myframe^{n-1,k\geq p} \arrow[d] \arrow[ddr] & \restrframe^{n,k\geq p} \arrow[d] \arrow[ddr] \arrow[l, dotted] & &
 frame^{n-1,p} \arrow[d] \arrow[ddr] & \restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \cohframe^{n+1,p} \arrow[d] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k\geq p}} \arrow[d, dotted] & &
   \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] &
   \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted] &
   \textcolor{red}{cohpainting^{n,k<p}} \arrow[l, dotted] \arrow[d, dotted] \\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & \textcolor{red}{cohframe^{n,k\geq p},X_{n-1}} \arrow[l, dotted] &
   \textcolor{red}{frame^{n-2,k<p}} &
   \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\end{tikzcd}
\end{scriptsize}

\item To assume $\cohpainting^{n+1,k}$ for $k \geq p$ to build
  $\cohframe^{n+1,k\geq p}$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{scriptsize}
\begin{tikzcd}
& & & frame^{n+1,p} \arrow[d] \arrow[ddr] \\
& & & painting^{n,p} \arrow[ddlll] \arrow[dddr] \arrow[d, dotted]\\
& & & frame^{n,p} \arrow[d] \arrow[ddr] & \restrframe^{n+1,k<p} \arrow[d] \arrow[ddr] \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\painting^{n-1,k\geq p} \arrow[d, dotted] \arrow[dddr] \arrow[dd, bend right=70] & & &
  painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] & \restrpainting^{n,k<p} \arrow[d, dotted] \arrow[ddlll] \arrow[dddll] \arrow[l, dotted] \arrow[uul, dotted]\\
\myframe^{n-1,k\geq p} \arrow[d] \arrow[ddr] & \restrframe^{n,k\geq p} \arrow[d] \arrow[ddr] \arrow[l, dotted] & &
 frame^{n-1,p} \arrow[d] \arrow[ddr] & \restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \cohframe^{n+1,p} \arrow[d] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] & \textcolor{red}{restrpainting^{n-1,k\geq p}} \arrow[d, dotted] & &
   \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] &
   \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted] &
   \textcolor{red}{cohpainting^{n,k<p}} \arrow[l, dotted] \arrow[d, dotted] \\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & \textcolor{red}{cohframe^{n,k\geq p},X_{n-1}} \arrow[l, dotted] &
   \textcolor{red}{frame^{n-2,k<p}} &
   \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\end{tikzcd}
\end{scriptsize}

\item To rely on $\restrpainting^{n,k}$ and $\cohframe^{n+1,k}$ for $k \geq
  p$ to also get $\restrpainting^{n+1,p}$ for all $p \leq n$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{scriptsize}
\begin{tikzcd}
& & & frame^{n+1,p} \arrow[d] \arrow[ddr] \\
& & & painting^{n,p} \arrow[ddlll] \arrow[dddr] \arrow[d, dotted]\\
& & & frame^{n,p} \arrow[d] \arrow[ddr] & \restrframe^{n+1,k<p} \arrow[d] \arrow[ddr] \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\painting^{n-1,k\geq p} \arrow[d, dotted] \arrow[dddr] \arrow[dd, bend right=70] & & &
  painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] & \restrpainting^{n,k<p} \arrow[d, dotted] \arrow[ddlll] \arrow[dddll] \arrow[l, dotted] \arrow[uul, dotted]\\
\myframe^{n-1,k\geq p} \arrow[d] \arrow[ddr] & \restrframe^{n,k\geq p} \arrow[d] \arrow[ddr] \arrow[l, dotted] & &
 frame^{n-1,p} \arrow[d] \arrow[ddr] & \restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \cohframe^{n+1,p} \arrow[d] \arrow[uul, dotted, shift right=2mm] \arrow[l, dotted]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] &
   \textcolor{red}{restrpainting^{n-1,k\geq p}} \arrow[l, dotted] \arrow[d, dotted] &
   \textcolor{red}{cohpainting^{n,k\geq p}} \arrow[l, dotted] \arrow[d, dotted] &
   \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] &
   \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted] &
   \textcolor{red}{cohpainting^{n,k<p}} \arrow[l, dotted] \arrow[d, dotted] \\
\textcolor{red}{frame^{n-2,k\geq p}} & \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] & \textcolor{red}{cohframe^{n,k\geq p},X_{n-1}} \arrow[l, dotted] &
   \textcolor{red}{frame^{n-2,k<p}} &
   \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\end{tikzcd}
\end{scriptsize}

\item To finally rely on $\cohpainting^{n,k}$ for $k \geq p$
to also get $\cohpainting^{n+1,p}$ for $p \leq n-1$:

$\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!$
\begin{scriptsize}
\begin{tikzcd}
& & & frame^{n+1,p} \arrow[d] \arrow[ddr] \\
& & & painting^{n,p} \arrow[ddlll] \arrow[dddr] \arrow[d, dotted]\\
& & & frame^{n,p} \arrow[d] \arrow[ddr] & \restrframe^{n+1,k<p} \arrow[d] \arrow[ddr] \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\painting^{n-1,k\geq p} \arrow[d, dotted] \arrow[dddr] \arrow[dd, bend right=70] & & &
 painting^{n-1,p} \arrow[ddlll] \arrow[dddll] \arrow[d, dotted] & \restrpainting^{n,k<p} \arrow[d, dotted] \arrow[ddlll] \arrow[dddll] \arrow[l, dotted] &
   \cohpainting^{n+1,p} \arrow[ddlll] \arrow[d, dotted] \arrow[l, dotted] \\
\myframe^{n-1,k\geq p} \arrow[d] \arrow[ddr] & \restrframe^{n,k\geq p} \arrow[d] \arrow[ddr] \arrow[l, dotted] & &
  frame^{n-1,p} \arrow[d] \arrow[ddr] & \restrframe^{n,k<p} \arrow[d] \arrow[ddr] \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \cohframe^{n+1,p} \arrow[d] \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm]\\
\textcolor{red}{painting^{n-2,k\geq p}} \arrow[d, dotted] &
   \textcolor{red}{restrpainting^{n-1,k\geq p}} \arrow[l, dotted] \arrow[d, dotted] &
   \textcolor{red}{cohpainting^{n,k\geq p}} \arrow[l, dotted] \arrow[d, dotted] &
   \textcolor{red}{painting^{n-2,k<p}} \arrow[d, dotted] &
   \textcolor{red}{restrpainting^{n-1,k<p}} \arrow[l, dotted] \arrow[d, dotted] &
   \textcolor{red}{cohpainting^{n,k<p}} \arrow[l, dotted] \arrow[d, dotted] \\
\textcolor{red}{frame^{n-2,k\geq p}} &
   \textcolor{red}{restrframe^{n-1,k\geq p},X_{n-2}} \arrow[l, dotted] &
   \textcolor{red}{cohframe^{n,k\geq p},X_{n-1}} \arrow[l, dotted] &
   \textcolor{red}{frame^{n-2,k<p}} &
   \textcolor{red}{restrframe^{n-1,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] &
   \textcolor{red}{cohframe^{n,k<p}} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm] \\
\end{tikzcd}
\end{scriptsize}
\end{enumerate}

Thus, shortly, assuming $\myframe^{n-2,p}$, $\painting^{n-2,p}$,
$\restrframe^{n-1,p}$, $\restrpainting^{n-1,p}$, $\cohframe^{n,p}$ and
$\cohpainting^{n,p}$ for all $p \leq n-2$ allows to define
$\myframe^{n-1,p}$, $\painting^{n-1,p}$, $\restrframe^{n,p}$,
$\restrpainting^{n,p}$, $\cohframe^{n+1,p}$ and $\cohpainting^{n+1,p}$
for all $p \leq n-1$, so the construction can be recursively
continued.

\end{document}