### 2025-12-23 paper/paper.tex
paper: better typesetting with some error fixes

\documentclass{msc}

\usepackage{amsmath, amssymb, mathrsfs, mathtools, wasysym, tikz, tikz-cd, lmodern, mathpazo, anyfontsize, xargs, environ, multirow, tabularx, caption, bookmark, booktabs, makecell, colortbl, minted, art.cls/colorpal, art.cls/ct, art.cls/sset, art.cls/lim, art.cls/joinargs}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[prefix=bonak]{art.cls/xkeymask}

% Use the patterns library to draw the cubes figure
\usetikzlibrary{patterns}

% Magic with xkeyval to go over the 9-argument limit of LaTeX
\makeatletter
\define@cmdkey[bonak]{X}{D}{(#1)}

\define@cmdkey[bonak]{frame}{D}{(#1)}
\define@cmdkey[bonak]{layer}{D}{(#1)}
\define@cmdkey[bonak]{layer}{d}{(#1)}
\define@cmdkey[bonak]{painting}{D}{(#1)}
\define@cmdkey[bonak]{painting}{E}{(#1)}
\define@cmdkey[bonak]{painting}{d}{(#1)}
\define@cmdkey[bonak]{fullframe}{D}{(#1)}

\define@cmdkey[bonak]{restrframe}{D}{(#1)}
\define@cmdkey[bonak]{restrframe}{d}{(#1)}
\define@cmdkey[bonak]{restrlayer}{D}{(#1)}
\define@cmdkey[bonak]{restrlayer}{d}{(#1)}
\define@cmdkey[bonak]{restrlayer}{l}{(#1)}
\define@cmdkey[bonak]{restrpainting}{D}{(#1)}
\define@cmdkey[bonak]{restrpainting}{E}{(#1)}
\define@cmdkey[bonak]{restrpainting}{d}{(#1)}
\define@cmdkey[bonak]{restrpainting}{c}{(#1)}

\define@cmdkey[bonak]{cohframe}{D}{(#1)}
\define@cmdkey[bonak]{cohframe}{d}{(#1)}
\define@cmdkey[bonak]{cohlayer}{D}{(#1)}
\define@cmdkey[bonak]{cohlayer}{d}{(#1)}
\define@cmdkey[bonak]{cohlayer}{l}{(#1)}
\define@cmdkey[bonak]{cohpainting}{D}{(#1)}
\define@cmdkey[bonak]{cohpainting}{E}{(#1)}
\define@cmdkey[bonak]{cohpainting}{d}{(#1)}
\define@cmdkey[bonak]{cohpainting}{c}{(#1)}
\makeatother

% Truncated sets, otherwise referred to as X: the general macro
\newcommandx{\X}[3][1,2,3]{
  \ensuremath{{\color{carolina}{\nu\textsf{Set}}}_{#1}^{#2}}
  \setkeys*[bonak]{X}{#3}
}

% Frame, layer, and painting: the general macro
\newcommandx{\prim}[6][2,3,4,5,6]{
  \ensuremath{\mathsf{\color{indian-yellow}{#1}}_{#2}^{
    \joinargs[#3][#4][#5]}}
  \setkeys*[bonak]{#1}{#6}
}

% Restriction for frame, layer, and painting: the general macro
\newcommandx{\restr}[8][2,3,4,5,6,7,8]{
  \ensuremath{\mathsf{\color{russian-green}{restr}}_{
    \joinargs[\mathsf{\color{indian-yellow}{#1}}][#2][#3][#4]}^{\joinargs[#5][#6][#7]}}
  \setkeys*[bonak]{restr#1}{#8}
}

\newcommandx{\depsp}[8][2,3,4,5,6,7,8]{
  \ensuremath{\mathsf{\color{russian-green}{deps}}_{
    \joinargs[\mathsf{\color{indian-yellow}{#1}}][#2][#3][#4]}^{\joinargs[#5][#6][#7]}}
}

\newcommandx{\depstypep}[8][2,3,4,5,6,7,8]{
  \ensuremath{\mathsf{\color{russian-green}{DEPS}}_{
    \joinargs[\mathsf{\color{indian-yellow}{#1}}][#2][#3][#4]}^{\joinargs[#5][#6][#7]}}
}

% Coherence conditions for frame, layer, and painting: the general macro

% Hack: we'd like \coh to take 10 arguments, with the 10th being passed to
% xkeyval, but this is not possible. The only possible compromise is to join
% #7 and #8 by hand in usage of \coh: it's now #7 or '#6,#7' in \cohfoo commands
\newcommandx{\coh}[9][2,3,4,5,6,7,8,9]{
  \ensuremath{\mathsf{\color{chestnut}{coh}}_{
    \joinargs[\mathsf{\color{indian-yellow}{#1}}][#2][#3][#4][#5][#6]}^{\joinargs[#7][#8]}}
  \setkeys*[bonak]{coh#1}{#9}
}

\newcommandx{\cohtwo}[9][2,3,4,5,6,7,8,9]{
  \ensuremath{\mathsf{\color{chestnut}{coh2}}_{
    \joinargs[\mathsf{\color{indian-yellow}{#1}}][#2][#3][#4][#5][#6]}^{\joinargs[#7][#8]}}
  \setkeys*[bonak]{coh#1}{#9}
}

% Simplicial and Cubical sets
\newcommand{\DeltaHat}{\ensuremath{\hat{\boldsymbol{\Delta}}}}
\newcommand{\DeltaPlus}{\ensuremath{\boldsymbol{\Delta}_+}}
\newcommand{\Cube}{\ensuremath{\boldsymbol{{\square}}}}
\newcommand{\CSet}{\ensuremath{\mathsf{Set}_{\boldsymbol{\square}}}}

% Universe, referred to as U
\newcommandx{\U}[1][1=]{\ensuremath{\mathsf{\color{spanish-blue}{HSet}}_{#1}}}
\newcommandx{\HGpd}[1][1=]{\ensuremath{\mathsf{\color{spanish-blue}{HGpd}}_{#1}}}
\newcommand{\Type}{\ensuremath{\mathsf{\color{spanish-blue}{Type}}}}
\newcommand{\SProp}{\ensuremath{\mathsf{\color{spanish-blue}{SProp}}}}

% Examples
\newtheorem{example}[therm]{Example}

% The unit and bool types
\newcommand{\unittype}{\ensuremath{\mathsf{unit}}}
\newcommand{\booltype}{\ensuremath{\mathsf{bool}}}
\newcommand{\unitpoint}{\ensuremath{\ast}}

% Definitional and extensional equality
\newcommand{\defeq}{\ensuremath{\triangleq}}
\newcommand{\refl}{\ensuremath{\mathsf{refl}}}

% Some abbreviations
\newcommand{\tl}{\ensuremath{\mathsf{tl}}}
\newcommand{\hd}{\ensuremath{\mathsf{hd}}}
\newcommand{\imp}{\rightarrow}
\newcommand{\overright}[1]{\overrightarrow{#1}}
\newcommand{\ap}{\mathsf{ap}\;}
\renewcommand{\D}{D}
\newcommand{\hdD}{D.1}
\newcommand{\tlD}{D.2}
\renewcommand{\d}{d}
\renewcommand{\E}{E}
\renewcommand{\l}{l}
\renewcommand{\c}{c}
\newcommand{\pair}[2]{#1, #2}
\newcommand{\Dom}{\textsf{Dom}}
\newcommand{\UIP}{\textsf{UIP}}

\newcommand{\udensdash}[1]{%
    \tikz[baseline=(todotted.base)]{
        \node[inner sep=1pt,outer sep=0pt] (todotted) {$#1$};
        \draw[densely dashed] (todotted.south west) -- (todotted.south east);
    }%
}%

% These are renewed in tables, but let's define them now for in-document
% typesetting
\newcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\newcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\newcommandx{\painting}[2][1,2]{\prim{painting}[][#1][#2][][]}
\newcommandx{\restrf}[4][1,2,3,4]{\restr{frame}[][#3][#4][#1][#2][][]}
\newcommandx{\restrl}[4][1,2,3,4]{\restr{layer}[][#3][#4][#1][#2][][]}
\newcommandx{\restrp}[4][1,2,3,4]{\restr{painting}[][#3][#4][#1][#2][][]}
\newcommandx{\cohf}{\coh{frame}[][][][][][][][]}
\newcommandx{\cohl}{\coh{layer}[][][][][][][][]}
\newcommandx{\cohp}{\coh{painting}[][][][][][][][]}
\newcommandx{\fullframe}[1][1]{\prim{fullframe}[][#1][][][]}

\newcommandx{\frametype}[2][1,2]{\prim{FRAME}[][#1][#2][][]}
\newcommandx{\layertype}[2][1,2]{\prim{LAYER}[][#1][#2][][]}
\newcommandx{\paintingtype}[2][1,2]{\prim{PAINTING}[][#1][#2][][]}
\newcommandx{\restrftype}[2][1,2]{\restr{FRAME}[][][][#1][#2][][]}
\newcommandx{\restrltype}[2][1,2]{\restr{LAYER}[][][][#1][#2][][]}
\newcommandx{\restrptype}[2][1,2]{\restr{PAINTING}[][][][#1][#2][][]}
\newcommandx{\cohftype}[2][1,2]{\coh{FRAME}[][][][][][#1][#2][]}
\newcommandx{\cohltype}[2][1,2]{\coh{LAYER}[][][][][][#1][#2][]}
\newcommandx{\cohptype}[2][1,2]{\coh{PAINTING}[][][][][][#1][#2][]}
\newcommandx{\cohttype}[2][1,2]{\cohtwo{FRAME}[][][][][][#1][#2][]}

\newcommandx{\deps}[2][1,2]{\depsp{}[][][][#1][#2][][]}
\newcommandx{\fulldeps}[1][1]{\depsp{full}[][][][#1][][][]}
\newcommandx{\depscohs}[2][1,2]{\depsp{cohs}[][][][#1][#2][][]}
\newcommandx{\fulldepscohs}[1][1]{\depsp{fullcohs}[][][][#1][][][]}
\newcommandx{\depscoht}[2][1,2]{\depsp{coht}[][][][#1][#2][][]}
\newcommandx{\fulldepscoht}[1][1]{\depsp{fullcoht}[][][][#1][][][]}
\newcommandx{\depstype}[2][1,2]{\depstypep{}[][][][#1][#2][][]}
\newcommandx{\fulldepstype}[1][1]{\depstypep{full}[][][][#1][][][]}
\newcommandx{\depscohstype}[2][1,2]{\depstypep{cohs}[][][][#1][#2][][]}
\newcommandx{\fulldepscohstype}[1][1]{\depstypep{fullcohs}[][][][#1][][][]}
\newcommandx{\depscohttype}[2][1,2]{\depstypep{coht}[][][][#1][#2][][]}
\newcommandx{\fulldepscohttype}[1][1]{\depstypep{fullcoht}[][][][#1][][][]}

\newcommandx{\coht}[7][1,2,3,4,5,6,7]{\cohtwo{frame}[#3][#4][#5][#6][#7][#1][#2][]}

% For typesetting type theory rules
\newcommand \seqr[3]
  {\shortstack{$#2$ \\ \mbox{}\\
                   \mbox{}\hrulefill\mbox{}\\ \mbox{}\\ $#3$} \raisebox{2ex}{$\;\;\mbox{$#1$}$}}

% Kerning for star
\newcommand{\kstar}{{\star}}

% Caption setup
\DeclareCaptionFormat{plain}{#1#3}
\captionsetup{font=footnotesize,labelfont=bf}

% The eqntable environment
\newcolumntype{Y}{>{\centering\arraybackslash}X}
\NewEnviron{eqntable}[1]{
  \fontsize{7.2}{9}\selectfont
  \begin{tabularx}{0.94\linewidth}{
    @{}
    >{$}l<{$}
    >{$}c<{$}
    >{$}c<{$}
    >{$}Y<{$}
    @{}}
    \toprule
    \BODY
    \bottomrule
  \end{tabularx}
  \captionof{table}{#1}
}

% \midrule color
\def\graymidrule{\arrayrulecolor{gray30}\midrule\arrayrulecolor{gray65}}

% A block in the eqntable environment
\NewDocumentCommand{\eqnline}{m m m m}{#1 & #2 & #3 & #4 \\}
\newcommandx*{\mc}[1]{\multicolumn{4}{c}{\emph{#1}} \\\\}

% \eqnarg is slightly more involved: output ( ... ) if not in mask,
% and { ... } if in mask
\newcommandx*{\eqnarg}[3]{\ifinmask[bonak]{#1}[#2]{\{#2:#3\}}{(#2:#3)}}

% Double-spacing
% \linespread{2}

\begin{document}
\title{The very dependent recursive structure of iterated parametricity in indexed form}
\begin{authgrp}
  \author{Hugo Herbelin}
  \affiliation{Université Paris Cité, Inria, CNRS, IRIF, Paris
    \email{Hugo.Herbelin@inria.fr}}
  \author{Ramkumar Ramachandra}
  \affiliation{Unaffiliated \email{r@artagnon.com}}
\end{authgrp}
\begin{abstract}
  In a previous paper, we gave equations describing the construction of semi-simplicial and semi-cubical sets via iterated parametricity. The construction was formalised in the Rocq (formerly, Coq) proof assistant building $n$-truncated sets using an induction on a large structure embedding equational reasoning.

  The current paper describes a new and lighter presentation of the construction where equational reasoning is replaced by definition. Being lighter, it can now be described in full details, closely following the corresponding new machine-checked formalisation.
\end{abstract}
\maketitle
\vspace{-2em} % Remove space after abstract

\section{Introduction}
In the previous paper~\citep{hr25}, we described the construction of augmented semi-simplicial and semi-cubical sets in \emph{indexed form} as a special case of iterated parametricity, meaning that instead of the following illustration of the cubical case (upto faces identities),

\begin{equation*}
  \begin{tikzcd}
    X_0: \U & X_1: \U \arrow[l, "\partial^L" description, shift left=2] \arrow[l, "\partial^R" description, shift right=2] & X_2: \U \arrow[l, "\partial^{L\kstar}" description, shift left=6] \arrow[l, "\partial^{R\kstar}" description, shift left=2] \arrow[l, "\partial^{\kstar L}" description, shift right=2] \arrow[l, "\partial^{\kstar R}" description, shift right=6] & \ldots
  \end{tikzcd}
\end{equation*}
we consider a family of families as shown below:
\begin{equation*}
  \begin{array}{lll}
    E_0 & :               \U                                                                          \\
    E_1 & :               E_0 \times X_0 \rightarrow  \U                                              \\
    E_2 & : \Pi a b c d.\,  E_1(a,b) \times E_1 (c,d) \times E_1(a,c) \times E_1 (b,d) \rightarrow \U \\
    \ldots
  \end{array}
\end{equation*}

% Abbreviated tables in this section
%
% The rule for implicit arguments for each table:
% show 1 explicit in first two parts
% show E + 1 explicit in the third part
\appendmask[bonak]{layer}[D]
\appendmask[bonak]{painting}[D]
\appendmask[bonak]{restrframe}[D]
\appendmask[bonak]{restrlayer}[D, d]
\appendmask[bonak]{restrpainting}[D, d]
\appendmask[bonak]{cohframe}[D]
\appendmask[bonak]{cohlayer}[D, d]
\appendmask[bonak]{cohpainting}[D, d]

% Truncated sets, otherwise referred to as X
% Keep all arguments
\newcommandx{\Xp}[1]{\X[#1][][]}
\newcommandx{\Xto}[3][3=]{\X[#1][<#2][#3]}
\newcommandx{\Xcomp}[3][3=]{\X[#1][=#2][#3]}
\newcommandx{\Xfrom}[3][3=]{\X[#1][\geq#2][#3]}

% Frame, layer, and painting
% Drop the universe letter
\renewcommandx{\framep}[5][1,2,3,4,5]{\prim{frame}[][#2][#3][#4][#5]}
\renewcommandx{\layer}[5][1,2,3,4,5]{\prim{layer}[][#2][#3][#4][#5]}
\renewcommandx{\painting}[5][1,2,3,4,5]{\prim{painting}[][#2][#3][#4][#5]}

% Restriction for frame, layer, and painting
% Drop the universe letter
\renewcommandx{\restrf}[7][1,2,3,4,5,6,7]{\restr{frame}[][#2][#3][#4][#5][#6][#7]}
\renewcommandx{\restrl}[7][1,2,3,4,5,6,7]{\restr{layer}[][#2][#3][#4][#5][#6][#7]}
\renewcommandx{\restrp}[7][1,2,3,4,5,6,7]{\restr{painting}[][#2][#3][#4][#5][#6][#7]}

% Coherence conditions for frame, layer, and painting
% Keep only ε, ω; n, p; proposition on n and p
\renewcommandx{\cohf}[9][1,2,3,4,5,6,7,8,9]{\coh{frame}[][#2][#3][][][][][#9]}
\renewcommandx{\cohl}[9][1,2,3,4,5,6,7,8,9]{\coh{layer}[][#2][#3][][][][][#9]}
\renewcommandx{\cohp}[9][1,2,3,4,5,6,7,8,9]{\coh{painting}[][#2][#3][][][][][#9]}

A $\nu$-set in indexed form is a sequence of families of $\U$, that is $\U[m]$ for some universe level $m$. We call such sequence a $\nu$-set at level $m$, whose type thus lives in $\U[m+1]$.

Table~\ref{tab:coind} describes the type of a $\nu$-set at level $m$ as a dependent stream of type families representing the limit of $n$-truncated $\nu$-sets: the recursive equation $\Xfrom{m}{n}{D} \,\defeq\, \Sigma R: \Xcomp{m}{n}[D=\D].\, \Xfrom{m}{n+1}[D=\pair{D}{R}]$ from the table formally corresponds to the stream $Stream_{\Sigma n.\, \Xto{m}{n},\, \lambda (n, \D).\, \Xcomp{m}{n}[D=\D],\,\lambda ((n, D),R).(n+1,(\pair{D}{R}))}(n, \D)$. Therefore, $\Xfrom{m}{n}$ denotes an infinite sequence $X_{n}, X_{n+1}, \ldots$ dependent on a $(<n)$-truncated $\nu$-set, $\Xto{m}{n}$, so that, when $n$ is $0$, it denotes a full $\nu$-set, written $\Xp{m}$. This is made possible because the $(<0)$-truncated $\nu$-set, $\Xto{m}\lbrack 0,0 \rbrack$, is degenerated: it is an empty family, and there is thus only one $(<0)$-truncated $\nu$-set, namely the canonical inhabitant $\kstar$ of $\unittype$.

The definition of the type of a $n$-truncated $\nu$-set is in turn described in table~\ref{tab:core}. In the infinite sequence of type families representing a $\nu$-set, the $n$-th component is a type dependent over a $\fullframe$. It is recursively defined in table~\ref{tab:frames}, using the auxiliary definitions of $\framep$, $\layer$ and $\painting$. A $\fullframe$ describes a boundary of a standard form (simplex, cube), which we decompose into $\layer$, and a $\painting$ corresponds to a filled frame. Notice that the type $\layer$ relies on an operator of frame restriction $\restrf$ which is defined in table~\ref{tab:faces}, and this restriction operator is in turn defined using auxiliary definitions $\restrl$ and $\restrp$.

Notably, the definition of $\restrl$ relies on an equality expressing the commutation of the composition of two $\restrf$. The proof of this commutation is worth being made explicit, which we do in table~\ref{tab:coh} using proof-term notations. The proof requires an induction on the dimension and on the structure of $\framep$, $\layer$, and $\painting$. This is what $\cohf$ does using auxiliary proofs $\cohl$ and $\cohp$. Even though it looks independent of the definitions from the other tables, $\cohf$ has to be proved mutually with the definitions of $\framep$, $\layer$, $\painting$, and their corresponding restrictions. More precisely, for a fixed $n$, the block of $\framep$, $\restrf$, and $\cohf$ has to be defined in one go by induction on $p$. Also, each of $\painting$, $\restrp$, and $\cohp$ is built by induction from $p$ to $n$. The $\painting$ block at $n$ relies on the $\framep$ block at $n$, but the converse dependency is only on lower $n$, so this is well-founded. Note that $\layer$, $\restrl$ and $\cohl$ are just abbreviations.

Note that for a fixed constant $n$, relying on the evaluation rules of type theory, the coherence conditions degenerate to a reflexivity proof, so that the construction builds an effective sequence of types not mentioning coherences anymore. \\

\def\lab{tab:coind}
\input{tab-coind.tex}

% For just the second table, keep the universe letter in fullframe
\renewcommandx{\fullframe}[3][1,2,3]{\prim{fullframe}[#1][#2][][][#3]}

\def\lab{tab:core}
\input{tab-core.tex}

\def\lab{tab:frames}
\input{tab-frames.tex}

% The third table mentions cohframe; keep ε, ω in this instance
\renewcommandx{\cohf}[9][1,2,3,4,5,6,7,8,9]{\coh{frame}[][#2][#3][][][#6,#7][#8][#9]}

\def\lab{tab:faces}
\input{tab-faces-ett.tex}

% Restriction for frame, layer, and painting
% Drop just the universe letter for the coh table
\renewcommandx{\restrf}[7][1,2,3,4,5,6,7]{\restr{frame}[][#2][#3][#4][#5][#6][#7]}
\renewcommandx{\restrl}[7][1,2,3,4,5,6,7]{\restr{layer}[][#2][#3][#4][#5][#6][#7]}
\renewcommandx{\restrp}[7][1,2,3,4,5,6,7]{\restr{painting}[][#2][#3][#4][#5][#6][#7]}

% Coherence conditions for frame, layer, and painting
\renewcommandx{\cohf}[9][1,2,3,4,5,6,7,8,9]{\coh{frame}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohl}[9][1,2,3,4,5,6,7,8,9]{\coh{layer}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohp}[9][1,2,3,4,5,6,7,8,9]{\coh{painting}[][#2][#3][#4][#5][#6,#7][#8][#9]}

\def\lab{tab:coh}
\input{tab-coh-ett.tex}

\renewcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\renewcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\renewcommandx{\painting}[2][1,2]{\prim{painting}[][#1][#2][][]}
\renewcommandx{\restrf}[4][1,2,3,4]{\restr{frame}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrl}[4][1,2,3,4]{\restr{layer}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrp}[4][1,2,3,4]{\restr{painting}[][#3][#4][#1][#2][][]}
\renewcommandx{\cohf}[6][1,2,3,4,5,6]{\coh{frame}[][#3][#4][#5][#6][#1][#2][]}
\renewcommandx{\cohl}[6][1,2,3,4,5,6]{\coh{layer}[][#3][#4][#5][#6][#1][#2][]}
\renewcommandx{\cohp}[6][1,2,3,4,5,6]{\coh{painting}[][#3][#4][#5][#6][#1][#2][]}

\section{Outline of the recursive structure of iterated parametricity construction}

In \citep{hr25}, the well-foundedness was justified by induction on $n$ and subinductions, one on $p$ to mutually build $\framep[n][p]$, $\restrf[n][p]$ and $\cohf[n][p]$, and three others to build each of $\painting[n][p]$, $\restrp[n][p]$ and $\cohp[n][p]$ by induction on $n-p$.

Since $\restrf[n][p]$ is referring to $\framep[n][p]$ and $\framep[n-1][p]$, and $\cohf[n][p]$ is referring to $\restrf[n][p]$ and $\restrf[n-1][p]$, thus also to $\framep[n][p]$, $\framep[n-1][p]$ and $\framep[n-2][p]$, we built all of these 6 components at the same time. But equations were also needed to state the properties of $\framep[n][p]$, $\framep[n-1][p]$ and $\restrf[n][p]$ so that we were building by induction on $n$, for any $D:\Xto{m}{n}$, the following block of information:

\begin{align*}
  \framep[n][p]      & : \U                                                                                                                 & \mbox{for all $p\leq n$}   \\
  \framep[n-1][p]    & : \U                                                                                                                 & \mbox{for all $p\leq n-1$} \\
  \framep[n-2][p]    & : \U                                                                                                                 & \mbox{for all $p\leq n-2$} \\
  \restrf[n][p]      & : \Pi q.\, \Pi \omega.\, \framep[n][p] \rightarrow \framep[n-1][p]                                                   & \mbox{for all $p\leq n$}   \\
  \restrf[n-1][p]    & : \Pi q.\, \Pi \omega.\, \framep[n-1][p] \rightarrow \framep[n-2][p]                                                 & \mbox{for all $p\leq n-1$} \\
  \cohf[n][p]        & : \Pi qr\epsilon\omega.\, \restrf[n-1][p][q][\epsilon] \circ \restrf[n][p][\omega][r] =                                                           \\
                     & \;\;\;\;\restrf[n-1][p][r][\omega] \circ \restrf[n][p][\epsilon][q+1]                                                & \mbox{for all $p\leq n$}   \\
  \mathsf{hypF0}     & : \framep[n][0]         = \unitpoint
  \\
  \mathsf{hypFS}     & : \framep[n][p+1]       = \Sigma d: \framep[n+1][p]. \Pi \omega.\, \painting[n][p] (\restrf[n+1][p](d))              & \mbox{for all $p\leq n-1$} \\
  \mathsf{hypPrevF0} & : \framep[n-1][0]                                                                                       = \unitpoint                              \\
  \mathsf{hypPrevFS} & : \framep[n-1][p+1] = \Sigma d: \framep[n][p]. \Pi \omega.\, \painting[n-1][p] (\restrf[n][p](d))                    & \mbox{for all $p\leq n-2$} \\
  \mathsf{hypRF0}    & : \restrf[n][0]        = \ldots \texttt{depending on hypF0}                                                                                       \\
  \mathsf{hypRFS}    & : \restrf[n][p+1]      = \ldots \texttt{depending on hypFS and hypPrevFS}                                            & \mbox{for all $p\leq n-1$} \\
  \ldots             & \;\;\;\;\texttt{fields for painting, restrf, cohp}
\end{align*}

In the current paper and associated formalisation, we switch the point
of view, take $\framep[n-2][p]$ $\painting[n-2][p]$, $\restrf[n-1][p]$
and $\restrp[n-1][p]$ as references and rename them respectively as
$\framep[n][p]$, $\painting[n][p]$, $\restrf[n][p]$ and
$\restrp[n][p]$. We replace the equational properties of $\framep[n][p]$,
$\framep[n-1][p]$, $\restrf[n][p]$, $\painting[n][p]$,
$\painting[n-1][p]$ and $\restrp[n][p]$ with definitions, renaming them
respectively to $\framep[n+2][p]$, $\framep[n+1][p]$,
$\restrf[n+1][p]$, $\painting[n+2][p]$, $\painting[n+1][p]$ and
$\restrp[n+1][p]$. This means that we build instead by induction a block of the form:

\begin{align*}
  \framep[n][p]   & :                                                                                                           \U                                                                                    & \mbox{for all $p\leq n$} \\
  \painting[n][p] & :                                                                                                           \framep[n][p] \rightarrow \U                                                          & \mbox{for all $p\leq n$} \\
  \restrf[n][p]   & :                                                                                                           \Pi q \leq n-p.\, \Pi \epsilon.\, \framep[n+1][p] \rightarrow \framep[n][p]           & \mbox{for all $p\leq n$} \\
  \restrp[n][p]   & :                                                                                                           \Pi q \leq n-p.\, \Pi \epsilon.\, \Pi d:\framep[n+1][p].\,                                                       \\
                  & \;\;\;\;\painting[n+1][p](d) \rightarrow \painting[n][p](\restrf[n][p](d))                                                                                                                        & \mbox{for all $p\leq n$} \\
  \cohf[n][p]     & :                                                                                                           \Pi qr\epsilon\omega.\, \restrf[n][p][q][\epsilon] \circ \restrf[n+1][p][\omega][r] =                            \\
                  & \;\;\;\;\restrf[n][p][\omega][r] \circ \restrf[n+1][p][q+1][\epsilon]                                                                                                                             & \mbox{for all $p\leq n$} \\
  \cohp[n][p]     & \Pi qr\epsilon\omega.\,\Pi d:\framep[n+2][p].\,                                                                                                                                                                              \\
                  & \;\;\;\;\restrp[n][p][q][\epsilon](\restrf[n+1][p][r][\omega](d)) \circ \restrp[n+1][p][r][\omega](d) =                                                                                                                      \\
                  & \;\;\;\;\restrp[n][p][r][\omega](\restrf[n+1][p][q+1][\epsilon](d)) \circ \restrp[n+1][p][q+1][\epsilon](d)                                                                                       & \mbox{for all $p\leq n$} \\
\end{align*}

Turning $\framep[n+2][p]$, $\framep[n+1][p]$ and $\restrf[n+1][p]$ into definitions involves however circularities. To start with, the definition of $\framep[n+1][p]$, which is by induction on $p$, depends on $\restrf[n][p-1]$ whose type depends in turn on $\framep[n+1][p-1]$. This means that $\framep[n+1][p]$ and the type of $\restrf[n][p]$ have to be mutually defined.

The definition of $\framep[n+1][p]$ also depends, via $\layer[n][p-1]$, on $\painting[n][p-1]$. Both via $\painting[n][p-1]$ and $\restrf[n][p-1]$, it depends also on $\framep[n][p-1]$ by typing dependency. This requires to assume $\framep[n][p-1]$ and $\painting[n][p-1]$ known in advance.
Moreover, the definition of $\framep[n+1][p]$ depends on $\framep[n+1][p-1]$, so all of $\framep[n][k]$, $\painting[n][k]$ and $\restrf[n][k]$ for $k< p$ need to be known, which we respectively abbreviate as $\framep[n][\lbrack 0,p-1 \rbrack]$, $\painting[n][\lbrack 0,p-1 \rbrack]$ and $\restrf[n][p-1]$.
Schematically, using dotted arrows for dependencies in the type and plain arrows for dependencies in the definition, this gives the following dependencies for $p\leq n+1$ fixed:
\begin{center}
  \begin{tikzcd}
    \framep[n+1][\lbrack 0,p \rbrack] \arrow[d] \arrow[ddr]\\
    \painting[n][\lbrack 0,p-1 \rbrack] \arrow[d, dotted] \\
    \framep[n][\lbrack 0,p-1 \rbrack] &
    \restrf[n][\lbrack 0,p-1 \rbrack] \arrow[l, dotted] \arrow[uul, dotted, "\mbox{$[0,p-1]$}"', near start, shift right=2mm, hook']\\
  \end{tikzcd}
\end{center}

In particular, the mutual dependency on one side of $\framep[n][\lbrack 0,p-1 \rbrack]$ within the type of $\restrf[n][\lbrack 0,p-1 \rbrack]$ and on the other side of $\restrf[n][\lbrack 0,p-1 \rbrack]$ within the definition of $\framep[n][\lbrack 0,p \rbrack]$ requires to mutually define the type of $\restrf[n][\lbrack 0,p-1 \rbrack]$ together with $\framep[n][p]$ (or $\framep[n][\lbrack 0,p \rbrack]$) as a \emph{function} from $\restrf[n][\lbrack 0,p-1 \rbrack]$.

So far, using this recursive definition of  $\framep[n][\lbrack 0,p-1 \rbrack]$ and $\restrf[n][\lbrack 0,p-1 \rbrack]$, we are able to specify the three first component of the exected dependent structure:
\begin{align*}
  \framep[n][p]             & :                        \U                                                & \mbox{for all $p\leq n$} \\
  \painting[n][p]           & :                        \framep[n][p] \rightarrow \U                      & \mbox{for all $p\leq n$} \\
  \restrf[n][p]             & :                        \Pi q \leq n-p.\, \Pi \epsilon.\, \framep[n+1][p]
  \left(\begin{array}{l}
            \framep[n][\lbrack 0,p-1 \rbrack]   \\
            \painting[n][\lbrack 0,p-1 \rbrack] \\
            \restrf[n][\lbrack 0,p-1 \rbrack]   \\
          \end{array}\right)
  \rightarrow \framep[n][p] & \mbox{for all $p\leq n$}                                                                              \\
\end{align*}
which can appropriately be reformulated using sequences for appropriate definitions of $\frametype[n][\lbrack 0,p \rbrack]$, $\paintingtype[n][\lbrack 0,p \rbrack]$ and $\restrftype[n][\lbrack 0,p \rbrack]$, since the dependencies of $\restrf[n][p]$ into $\restrf[n][\lbrack 0,p-1 \rbrack]$ makes it a very dependent function~\citep{Hickey96}:
\begin{align*}
  \framep[n][\lbrack 0,p \rbrack]   & : \frametype[n][\lbrack 0,p \rbrack]                                     \\
  \painting[n][\lbrack 0,p \rbrack] & : \paintingtype[n][\lbrack 0,p \rbrack](\framep[n][\lbrack 0,p \rbrack]) \\
  \restrf[n][\lbrack 0,p \rbrack]   & : \restrftype[n][\lbrack 0,p \rbrack]
  \left(\begin{array}{l}
            \framep[n][\lbrack 0,p-1 \rbrack]   \\
            \painting[n][\lbrack 0,p-1 \rbrack] \\
            \restrf[n][\lbrack 0,p-1 \rbrack]   \\
          \end{array}\right)                                                                    \\
\end{align*}

The type of sequences such as $\restrf[n][\lbrack 0,p \rbrack]$ will be represented by iterated $\Sigma$-types since each $\restrf[n][p]$ may depend on the previous components of the sequence. In the case of $\framep$ and $\painting$, there will be however no dependency, so an ordinary product can be used to represent $\framep[n][\lbrack 0,p \rbrack]$ and $\painting[n][\lbrack 0,p \rbrack]$. Later on, we will write $\framep[n][\lbrack 0,p-1 \rbrack]$ for the immediate subsequence of length $p$ of $\framep[n][\lbrack 0,p \rbrack]$ and $\framep[n][p]$ for its last component, and similarly for $\painting[n][\lbrack 0,p \rbrack]$, $\restrf[n][\lbrack 0,p \rbrack]$, etc.

\section{The construction}

\subsection{Specifying frames}

Assuming $\framep[n][\lbrack 0,p-1 \rbrack]$ for $p \leq n+1$ only requires to define the type of frames, which are just finite products of $\U$:

\begin{equation*}
  \begin{array}{llcl}
    \frametype[n][\lbrack 0,p-1 \rbrack] &  & :      & \U                                               \\
    \frametype[n]                        &  & \defeq & \unittype                                        \\
    \frametype[n][\lbrack 0,p \rbrack]   &  & \defeq & \frametype[n+1][\lbrack 0,p-1 \rbrack] \times \U \\
  \end{array}
\end{equation*}

\subsection{Specifying paintings}

The types $\paintingtype[n][\lbrack 0,p-1 \rbrack]$ of $\painting[n][\lbrack 0,p-1 \rbrack]$ for $p \leq n+1$ are type families over frames, so $\framep[n][\lbrack 0,p-1 \rbrack]$ needs to be assumed beforehand:

\begin{equation*}
  \begin{array}{llcl}
    \paintingtype[n][\lbrack 0,p-1 \rbrack] & (\framep[n][\lbrack 0,p \rbrack]:\frametype[n][\lbrack 0,p-1 \rbrack]) & :      & \U                                                                                                                   \\
    \paintingtype[n]                        & \unitpoint                                                             & \defeq & \unittype                                                                                                            \\
    \paintingtype[n][\lbrack 0,p \rbrack]   & (\framep[n][\lbrack 0,p \rbrack])                                      & \defeq & \paintingtype[n+1][\lbrack 0,p-1 \rbrack](\framep[n+1][\lbrack 0,p-1 \rbrack]) \times (\framep[n][p] \rightarrow \U) \\
  \end{array}
\end{equation*}

\subsection{Defining frames and specifying frame restrictions}

Assuming $\framep[n][\lbrack 0,p-1 \rbrack]$ and $\painting[n][\lbrack 0,p-1 \rbrack]$ for $p \leq n+1$, we can
mutually define the types $\restrftype[n][\lbrack 0,p-1 \rbrack]$ of frame
restrictions together with $\framep[n+1][\lbrack 0,p-1 \rbrack]$ as a function from
$\restrf[n][\lbrack 0,p-1 \rbrack]$, introducing also abbreviations for the types
and definitions of layers, as well as for the combination of sequences
of $\framep[n][\lbrack 0,p-1 \rbrack]$, $\painting[n][\lbrack 0,p-1 \rbrack]$ and
$\restrf[n][\lbrack 0,p-1 \rbrack]$ in one block, which we call
$\depstype[n][0,...p-1]$:

\begin{equation*}
  \begin{array}{llcl}
    \depstype[n][\lbrack 0,p-1 \rbrack]                                                                               &        & \defeq                               &
    \left(\begin{array}{lll}
              \framep[n][\lbrack 0,p-1 \rbrack]   & : & \frametype[n][\lbrack 0,p-1 \rbrack]    \\
              \painting[n][\lbrack 0,p-1 \rbrack] & : & \paintingtype[n][\lbrack 0,p-1 \rbrack]
              (\framep[n][\lbrack 0,p-1 \rbrack])                                               \\
              \restrf[n][\lbrack 0,p-1 \rbrack]   & : & \restrftype[n][\lbrack 0,p-1 \rbrack]
              \left(\begin{array}{l}
                  \framep[n][\lbrack 0,p-1 \rbrack]   \\
                  \painting[n][\lbrack 0,p-1 \rbrack] \\
                \end{array}\right)                                         \\
            \end{array}\right)                                                                             \\
    \framep[n+1][\lbrack 0,p \rbrack]                                                                                 &
    (\deps[n][\lbrack 0,p-1 \rbrack]:\depstype[n][\lbrack 0,p-1 \rbrack])                                             & :      & \frametype[n+1][\lbrack 0,p \rbrack]   \\
    \framep[n+1][\lbrack 0,0 \rbrack]                                                                                 &
    \unitpoint                                                                                                        & \defeq & (\unitpoint,\unittype)                 \\
    \framep[n+1][\lbrack 0,p+1 \rbrack]                                                                               &
    \deps[n][\lbrack 0,p \rbrack]                                                                                     & \defeq &
    \left(\begin{array}{l}
              \framep[n+1][\lbrack 0,p \rbrack]
              (\deps[n][\lbrack 0,p-1 \rbrack]), \\
              \Sigma d:\framep[n+1][p]
              (\deps[n][\lbrack 0,p-1 \rbrack]).
              \, \layer[n][p]
              (\deps[n][\lbrack 0,p \rbrack])
              (d)                                \\
            \end{array}\right)                                                                                                                            \\
    \restrftype[n][\lbrack 0,p-1 \rbrack]                                                                             &
    \left(\begin{array}{l}
              \framep[n][\lbrack 0,p-1 \rbrack]:\frametype[n][\lbrack 0,p-1 \rbrack]                                         \\
              \painting[n][\lbrack 0,p-1 \rbrack]:\paintingtype[n][\lbrack 0,p-1 \rbrack](\framep[n][\lbrack 0,p-1 \rbrack]) \\
            \end{array}\right) & :      & \U                                                \\
    \restrftype[n]                                                                                                    &
    \left(\begin{array}{l}
              \unitpoint \\
              \unitpoint \\
            \end{array}\right)                                                                                            & \defeq &
    \unittype                                                                                                                                                           \\
    \restrftype[n][\lbrack 0,p \rbrack]                                                                               &
    \left(\begin{array}{l}
              \framep[n][\lbrack 0,p \rbrack]   \\
              \painting[n][\lbrack 0,p \rbrack] \\
            \end{array}\right)                                                                              & \defeq &
    \begin{array}{l}
      \Sigma \restrf[n][\lbrack 0,p-1 \rbrack]
      :\restrftype[n][\lbrack 0,p-1 \rbrack]
      \left(\begin{array}{l}
                \framep[n+1][\lbrack 0,p-1 \rbrack]   \\
                \painting[n+1][\lbrack 0,p-1 \rbrack] \\
              \end{array}\right).         \\
      \Pi q \leq n-p.\,\Pi \epsilon.\,\framep[n+1][p]
      \left(\begin{array}{l}
                \framep[n][\lbrack 0,p-1 \rbrack]   \\
                \painting[n][\lbrack 0,p-1 \rbrack] \\
                \restrf[n][\lbrack 0,p-1 \rbrack]   \\
              \end{array}\right) \rightarrow  \framep[n][p] \\
    \end{array}                                                                                                  \\
    \layertype[n][p]                                                                                                  &
    (\deps[n][\lbrack 0,p \rbrack]:\depstype[n][\lbrack 0,p \rbrack])
                                                                                                                      & :      & \U                                     \\
    \layertype[n][p]                                                                                                  &
    \deps[n][\lbrack 0,p \rbrack]
                                                                                                                      & \defeq & \framep[n+1][p]
    (\deps[n][\lbrack 0,p-1 \rbrack])
    \rightarrow \U                                                                                                                                                      \\
    \layer[n][p]                                                                                                      &
    (\deps[n][\lbrack 0,p \rbrack]:\depstype[n][\lbrack 0,p \rbrack])
                                                                                                                      & :      & \layertype[n][p]
    (\deps[n][\lbrack 0,p \rbrack])
    \\
    \layer[n][p]                                                                                                      &
    \deps[n][\lbrack 0,p \rbrack]
                                                                                                                      & \defeq &
    d \mapsto \Pi\epsilon.\,\painting[n][p](\deps.\restrf[n][p][0][\epsilon](d))                                                                                        \\
  \end{array}
\end{equation*}

Note that we write $\deps.\framep$, $\deps.\painting$, $\deps.\restrf$ for the
projections of a $\deps$. Also, by abuse of notation, we write $\unitpoint$ for
the only inhabitant of $\depstype[n]$.

\subsection{Defining paintings}

Defining $\painting[n+1][p]$ for $p \leq n+1$, in addition to the type dependency in $\framep[n+1][p]$, requires the definition of $\painting[n][p-1]$, thus also $\framep[n][p-1]$ by typing dependency, as well as $\restrf[n][p-1]$. Since $\painting[n+1][p]$ also depends on $\painting[n+1][p+1]$ when $p<n+1$, these dependencies are actually between $\painting[n+1][\lbrack p,n+1 \rbrack]$, $\painting[n][\lbrack p,n \rbrack]$, $\framep[n+1][\lbrack p,n+1 \rbrack]$, $\framep[n][\lbrack p,n \rbrack]$ and $\restrf[n][\lbrack p,n \rbrack]$. Also, for the case $p=n$, this requires the assumption of an inhabitant $X_{n}$ of $\nu$Set$^{=n}$. Finally, $\framep[n+1][\lbrack p,n+1 \rbrack]$, $\framep[n+1][\lbrack p,n \rbrack]$ and $\restrf[n][\lbrack p,n \rbrack]$ respectively depend on the earlier values $\framep[n+1][\lbrack 0,p-1 \rbrack]$, $\framep[n][\lbrack 0,p-1 \rbrack]$ and $\restrf[n][\lbrack 0,p-1 \rbrack]$.  Below, we use dashed underline and dashed arrows for dependencies in subconstructions, here in the construction of $\framep$ at level $n$ and $\restrftype$ at level $n$:
\begin{center}
  \begin{tikzcd}
    & & \painting[n+1][\lbrack p,n+1 \rbrack] \arrow[dd, bend right=70] \arrow[dddr] \arrow[d, dotted]\\
    \udensdash{\framep[n+1][\lbrack 0,p-1 \rbrack]} \arrow[d, dashed] \arrow[ddr, dashed] & &
    \udensdash{\framep[n+1][\lbrack p,n+1 \rbrack]} \arrow[ll, dotted] \arrow[d, dashed] \arrow[ddr, dashed, shift right=2mm] \\
    \painting[n][\lbrack 0,p-1 \rbrack] \arrow[d, dotted] & &
    \painting[n][\lbrack p,n-1 \rbrack] \arrow[ll, dotted] \arrow[d, dotted] \\
    \framep[n][\lbrack 0,p-1 \rbrack] &
    \udensdash{\restrf[n][\lbrack 0,p-1 \rbrack]} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm, hook'] &
    \framep[n][\lbrack p,n \rbrack] \arrow[ll, dotted, bend left=15] &
    \udensdash{\restrf[n+1][\lbrack p,n \rbrack]},X_n \arrow[ll, dotted, bend left=15] \arrow[l, dotted] \arrow[uul, dotted, hook']\\
  \end{tikzcd}
\end{center}

Let us introduce another abbreviation to combine a full $\deps[n][\lbrack 0,n \rbrack]:\depstype[n][\lbrack 0,n \rbrack]$ with a family $E:\framep[n+1][n+1](\deps[n][\lbrack 0,n \rbrack])\rightarrow \U$:
\begin{equation*}
  \fulldepstype[n] \defeq
  \left(\begin{array}{lcl}
      \deps[n][\lbrack 0,n \rbrack] & : & \depstype[n][\lbrack 0,n \rbrack]                               \\
      E                             & : & \framep[n+1][n+1](\deps[n][\lbrack 0,n \rbrack]) \rightarrow \U \\
    \end{array}\right)
\end{equation*}

TODO: say the convention on projection in the Introduction

Also, when $n$ is 0, we write $E$ to denote an element $(\unitpoint, E)$ of $\fulldepstype[0]$.

We can now formally define $\painting[n+1][p]$ for $p \leq n+1$ by recursion on $n-p$:
\begin{equation*}
  \begin{array}{llcl}
    \painting[n+1][p]               &
    (\fulldeps[n]:\fulldepstype[n]) & :      & \paintingtype[n+1][p]      \\
    \painting[n+1][n+1]             &
    \fulldeps[n]                    & \defeq & \fulldeps.E                \\
    \painting[n+1][p \leq n]        &
    \fulldeps[n]                    & \defeq & d \mapsto
    \left(\begin{array}{l}
              \Sigma l:\layer[n][p](\fulldeps.\deps[n][\lbrack 0,p \rbrack]). \\
              \painting[n+1][p+1]
              (\fulldeps[n])(d,l)                                             \\
            \end{array}\right) \\
  \end{array}
\end{equation*}

From which we can deduce $\painting[n+1][\lbrack 0,p-1 \rbrack]$ for $p \leq n+2$ by a second recursion, this time on $p$:

\begin{equation*}
  \begin{array}{llcl}
    \painting[n+1][\lbrack 0,p-1 \rbrack] &
    (\fulldeps[n]:\fulldepstype[n])       & :      & \paintingtype[n+1][\lbrack 0,p-1 \rbrack](\framep[n+1][\lbrack 0,p-1 \rbrack](\fulldeps.\deps[n][\lbrack 0,p-1 \rbrack])) \\
    \painting[n+1]                        &
    \fulldeps[n]                          & \defeq & \unittype                                                                                                                 \\
    \painting[n+1][\lbrack 0,p \rbrack]   &
    \fulldeps[n]
                                          & \defeq &
    (\painting[n+1][\lbrack 0,p-1 \rbrack](\fulldeps[n-1]),\painting[n+1][p](\fulldeps[n]))                                                                                    \\
  \end{array}
\end{equation*}

\subsection{Specifying painting restrictions}

At this stage, we are able to define $\framep[n+1][\lbrack 0,n+1 \rbrack]$, $\painting[n+1][\lbrack 0,n+1 \rbrack]$ and $\restrftype[n][\lbrack 0,n \rbrack]$ from $\framep[n][\lbrack 0,n \rbrack]$, $\painting[n][\lbrack 0,n \rbrack]$ and $\restrf[n][\lbrack 0,n \rbrack]$, that is from $\deps[n][\lbrack 0,n \rbrack]$. To go further, we need to define the sequence $\restrptype[n][\lbrack 0,p-1 \rbrack]$ of types of painting restrictions, for $p \leq n+1$:

\begin{equation*}
  \begin{array}{llcl}
    \restrptype[n][\lbrack 0,p-1 \rbrack] &
    (\fulldeps[n]:\fulldepstype[n])       & :      & \U                                                         \\
    \restrptype[n]                        & E      & \defeq                                         & \unittype \\
    \restrptype[n][\lbrack 0,p \rbrack]   &
    (\fulldeps[n])
                                          & \defeq & \restrptype[n+1][\lbrack 0,p-1 \rbrack] \times             \\
                                          &        &                                                & \left(
    \begin{array}{l}
        \Pi q \leq n-p.\Pi \epsilon.\Pi d:\framep[n+1][p](\fulldeps.\deps[n][\lbrack 0,p-1 \rbrack]). \\
        \painting[n+1][p]
        (\fulldeps[n])(d) \rightarrow                                                                 \\ \fulldeps.\deps.\painting[n][p](\fulldeps.\deps.\restrf[n][p][q][\epsilon](d))\end{array}
    \right)                                                                                                     \\
  \end{array}
\end{equation*}

\subsection{Defining frame restrictions and specifying frame coherence laws}

Thanks to $\restrptype[n][\lbrack 0,p-1 \rbrack]$ for $p \leq n+1$, we can mutually define $\restrf[n+1][\lbrack 0,p \rbrack]$ and $\cohftype[n][\lbrack 0,p-1 \rbrack]$ from $\framep[n][\lbrack 0,n \rbrack]$, $\painting[n][\lbrack 0,n \rbrack]$, $\restrf[n][\lbrack 0,n \rbrack]$, $\restrptype[n][\lbrack 0,p-1 \rbrack]$ and some $E:\framep[n+1][n+1](\deps[n][\lbrack 0,n \rbrack]) \rightarrow \U$, abbreviating these dependencies and the dependency in a $\cohf[n][\lbrack 0,p-1 \rbrack]$ as $\depscohstype[n][\lbrack 0,p-1 \rbrack]$, and using also abbreviations in passing for $\restrl$ and $\restrltype$. It is also convenient to build $\deps[n+1][\lbrack 0,p \rbrack]$ from $\depscohs[n][\lbrack 0,p-1 \rbrack]$, so that we can now build $\framep[n+2][p]$ one step further. Thus, we also include this abbreviation in the definition:
\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \depscohstype[n][\lbrack 0,p-1 \rbrack]                                                  &                                     & \defeq                                                         &
    \left(\begin{array}{lcl}
              \fulldeps[n]                    & : & \fulldepstype[n]                                    \\
              \restrp[n][0,...p-1]            & : & \restrptype[n][\lbrack 0,p-1 \rbrack](\fulldeps[n]) \\
              \cohf[n][\lbrack 0,p-1 \rbrack] & : & \cohftype[n][\lbrack 0,p-1 \rbrack]
              \left(\begin{array}{l}
                  \fulldeps[n]                      \\
                  \restrp[n][\lbrack 0,p-1 \rbrack] \\
                \end{array}\right)                                                   \\
            \end{array}\right)                                                                                                     \\
    \overline{\deps}                                                                         &
    (\depscohs[n][\lbrack 0,p-1 \rbrack]:\depscohstype[n][\lbrack 0,p-1 \rbrack])            & :                                   & \depstype[n+1][\lbrack 0,p \rbrack]                                \\
    \overline{\deps}                                                                         & \depscohs[n][\lbrack 0,p-1 \rbrack] & \defeq                                                         &
    \left(\begin{array}{l}
              \framep[n+1][\lbrack 0,p \rbrack](\depscohs.\fulldeps.\deps[n][\lbrack 0,p-1 \rbrack]) \\
              \painting[n+1][\lbrack 0,p \rbrack](\depscohs.\fulldeps[n])                            \\
              \restrf[n+1][\lbrack 0,p \rbrack](\depscohs[n][\lbrack 0,p-1 \rbrack])                 \\
            \end{array}\right)                                                                                                        \\
    \restrf[n+1][\lbrack 0,p \rbrack]                                                        &
    (\depscohs[n][\lbrack 0,p-1 \rbrack]:\depscohstype[n][\lbrack 0,p-1 \rbrack])            & :                                   &
    \restrftype[n+1][\lbrack 0,p \rbrack]
    \left(\begin{array}{l}
              \framep[n+1][\lbrack 0,p \rbrack](\depscohs.\fulldeps.\deps[n][\lbrack 0,p \rbrack]) \\
              \painting[n+1][\lbrack 0,p \rbrack](\depscohs.\fulldeps[n])                          \\
            \end{array}\right)                                                                                                          \\
    \restrf[n+1][\lbrack 0,0 \rbrack]                                                        &
    E
                                                                                             & \defeq                              & (\unitpoint,\lambda q. \lambda\epsilon. \lambda \_.\unitpoint)     \\
    \restrf[n+1][\lbrack 0,p+1 \rbrack]                                                      &
    (\depscohs[n][\lbrack 0,p \rbrack])                                                      & \defeq                              &
    \left(\begin{array}{l}
              \restrf[n+1][\lbrack 0,p \rbrack][q][\epsilon]
              (\depscohs[n][\lbrack 0,p-1 \rbrack]),           \\
              \lambda q.\lambda \epsilon.\lambda (d,l).        \\
              \left(\restrf[n][p][q][\epsilon]
                (\depscohs[n][\lbrack 0,p-1 \rbrack])(d), \restrl[n][p][q][\epsilon]
              (\depscohs[n][\lbrack 0,p \rbrack])(d)(l)\right) \\
            \end{array} \right)                                                                                                                                 \\
    \cohftype[n][\lbrack 0,p-1 \rbrack]                                                      &
    \left(\begin{array}{l}
              \fulldeps[n]:\fulldepstype[n]                                                         \\
              \restrp[n][\lbrack 0,p-1 \rbrack]:\restrptype[n][\lbrack 0,p-1 \rbrack](\fulldeps[n]) \\
            \end{array}\right) & :                                   & \U                                                                                                         \\
    \cohftype[n]                                                                             &
    \left(\begin{array}{l}
              E          \\
              \unitpoint \\
            \end{array}\right)                                                                   & \defeq                              &
    \unittype                                                                                                                                                                                           \\
    \cohftype[n][\lbrack 0,p \rbrack]                                                        &
    \left(\begin{array}{l}
              \fulldeps[n]                    \\
              \restrp[n][\lbrack 0,p \rbrack] \\
            \end{array}\right)                                                       & \defeq                              &
    \left\{\begin{array}{l}
             \Sigma \cohf[n][\lbrack 0,p-1 \rbrack][q][r][\epsilon][\omega]
             :\cohftype[n][\lbrack 0,p-1 \rbrack]
             \left(\begin{array}{l}
                \fulldeps[n]                      \\
                \restrp[n][\lbrack 0,p-1 \rbrack] \\
              \end{array}\right).                  \\
             \Pi d:\framep[n+2][p]
             (\overline{\deps}\left(\begin{array}{l}
                                 \fulldeps[n]                      \\
                                 \restrp[n][\lbrack 0,p-1 \rbrack] \\
                                 \cohf[n][\lbrack 0,p \rbrack]
                               \end{array}\right))          . \\
             \\
             \fulldeps.\deps.\restrf[n][p][q][\epsilon](\restrf[n+1][p][r][\omega]
             \left(\begin{array}{l}
                \fulldeps[n]                      \\
                \restrp[n][\lbrack 0,p-1 \rbrack] \\
                \cohf[n][\lbrack 0,p \rbrack]
              \end{array}\right)(d)) =                  \\
             \fulldeps.\deps.\restrf[n][p][r][\omega](\restrf[n+1][p][q+1][\epsilon]
             \left(\begin{array}{l}
                \fulldeps[n]                      \\
                \restrp[n][\lbrack 0,p-1 \rbrack] \\
                \cohf[n][\lbrack 0,p \rbrack]
              \end{array}\right)(d))                  \\
           \end{array}\right.                                                                                                                            \\
    \restrltype[n][p]                                                                        &
    (\depscohs[n][\lbrack 0,p-1 \rbrack]:\depscohstype[n][\lbrack 0,p-1 \rbrack])
                                                                                             & :                                   & \U                                                                 \\
    \restrltype[n][p]                                                                        &
    \depscohs[n][\lbrack 0,p-1 \rbrack]
                                                                                             & \defeq                              &
    \left\{\begin{array}{l}
             \Pi q \leq n-p. \Pi \epsilon. \Pi d: \framep[n+2][p](\overline{\deps}(\depscohs[n][\lbrack 0,p-1 \rbrack])).\, \\
             \layer[n+1][p](\overline{\deps}(\depscohs[n][\lbrack 0,p-1 \rbrack])(d) \rightarrow                            \\
             \layer[n][p](\fulldeps.\deps[n][\lbrack 0,p \rbrack])(\restrf[n][p]
             (\depscohs[n][\lbrack 0,p-1 \rbrack])(d))
           \end{array}\right.                                                                               \\
    \restrl[n][p]                                                                            &
    (\depscohs[n][\lbrack 0,p \rbrack]:\depscohstype[n][\lbrack 0,p \rbrack])                & :                                   &
    \restrltype[n][p] (\depscohs[n][\lbrack 0,p \rbrack])                                                                                                                                               \\
    \restrl[n][p][q][\epsilon]                                                               &
    \depscohs[n][\lbrack 0,p \rbrack]                                                        & \defeq                              &
    \lambda d.\lambda l.\lambda \omega\overrightarrow{\cohf[n][p][q][0][\epsilon][\omega](d)}{\restrp[n][p][q][\epsilon](\restrf[n][p][q][\epsilon](\depscohs[n][\lbrack 0,p-1 \rbrack])(d))(l_\omega)} \\
  \end{array}
\end{equation*}

By abuse of notation, we also write $E$ for an inhabitant of $\depscohstype[n]$ (standing for a non trivial inhabitant of $\fulldeps[n]$).

We need also the notion of full dependencies at the level of coherence conditions, that combines a full $\depscohs[n][\lbrack 0,n \rbrack]$ with a set $E:\framep[n+2][n+2](\overline{\deps}(\depscohs[n][\lbrack 0,n \rbrack])) \rightarrow \U$:
\begin{equation*}
  \fulldepscohstype[n] \defeq
  \left(\begin{array}{lcl}
      \depscohs[n][\lbrack 0,n-1 \rbrack] & : & \depscohstype[n][\lbrack 0,n \rbrack]                                                \\
      E                                   & : & \framep[n+2][n+2](\overline{\deps}(\depscohs[n][\lbrack 0,n \rbrack]))\rightarrow \U \\
    \end{array}\right)
\end{equation*}
TODO: say the convention on projection in the Introduction

We can then extend $\overline{\deps}$ to build $\fulldeps$ from $\fulldepscohs$:
\begin{align*}
  \overline{\fulldeps}(\fulldepscohs[n]:\fulldepscohstype[n]) & :     \fulldepstype[n+1]                                                       \\
  \overline{\fulldeps}(\fulldepscohs[n])                      & \defeq \left(\begin{array}{l}
                                                                                 \overline{\deps}(\fulldepscohs.\depscohs[n][\lbrack 0,n \rbrack]) \\
                                                                                 \fulldepscohs.E                                                   \\
                                                                               \end{array}\right) \\
\end{align*}

Also, when $n$ is 0, we write $(E,E')$ to denote an element $(E, E')$ of $\fulldepscohstype[0]$.

\subsection{Defining painting restrictions}

Armed with $\restrf$, $\cohftype$ and these abbreviations, it is now possible to define $\restrp[n+1][p]$ for $p \leq n+1$ by recursion on $n-p$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \restrp[n+1][p][q][\epsilon]            &
    (\fulldepscohs[n]:\fulldepscohstype[n]) & :      &
    \left(
    \begin{array}{l}
        \Pi d:\framep[n+2][p]
        (\overline{\deps}(\fulldepscohs.\depscohs[n][\lbrack 0,p-1 \rbrack])). \\
        \painting[n+2][p]
        (\overline{\fulldeps}(\fulldepscohs[n]))(d) \rightarrow                \\
        \painting[n+1][p](\fulldepscohs.\depscohs.\fulldeps[n])
        (\restrf[n+1][p](\fulldepscohs.\depscohs[n][\lbrack 0,p-1 \rbrack])(d))\end{array}\right)                        \\
    \restrp[n+1][n+1][q][\epsilon]          &
    (\fulldepscohs[n])                      & \defeq & \lambda d. \lambda (l,c). l\epsilon   \\
    \restrp[n+1][p \leq n][q][\epsilon]     &
    (\fulldepscohs[n])                      & \defeq & \lambda d. \lambda (l,c).
    \left(\begin{array}{l}
              \restrl[n][p][q][\epsilon](\fulldepscohs.\depscohs[n][\lbrack 0,p \rbrack])(d)(l), \\
              \restrp[n+1][p+1][q][\epsilon](\fulldepscohs[n])(d,l)(c)                           \\
            \end{array}\right) \\
  \end{array}
\end{equation*}

From which we can deduce $\restrp[n+1][\lbrack 0,p-1 \rbrack]$ for $p \leq n+2$ again by a second recursion, this time on $p$:

\begin{equation*}
  \begin{array}{llcl}
    \restrp[n+1][\lbrack 0,p-1 \rbrack]     &
    (\fulldepscohs[n]:\fulldepscohstype[n]) & :      & \restrptype[n+1][\lbrack 0,p-1 \rbrack](\fulldepscohs.\depscohs.\fulldeps[n])) \\
    \restrp[n+1]                            &
    \fulldepscohs[n]                        & \defeq & \unittype                                                                      \\
    \restrp[n+1][\lbrack 0,p \rbrack]       &
    \fulldepscohs[n]
                                            & \defeq &
    (\restrp[n+1][\lbrack 0,p-1 \rbrack](\fulldepscohs[n]),\restrp[n+1][p](\fulldepscohs[n]))                                         \\
  \end{array}
\end{equation*}

\subsection{Specifying painting coherence laws}

At this stage, we are then able to define $\restrf[n+1][\lbrack 0,p \rbrack]$, $\restrp[n+1][\lbrack 0,p \rbrack]$ and $\cohftype[n][\lbrack 0,p-1 \rbrack]$ for $p \leq n+1$
from $\depscohs[n][\lbrack 0,p-1 \rbrack]$ at the previous level. To go further, we need to define the
sequence $\cohptype[n][\lbrack 0,p-1 \rbrack]$ of types of painting coherence laws:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \cohptype[n][\lbrack 0,p-1 \rbrack] & (\fulldepscohs[n]:\fulldepscohstype[n]) & :      & \U                                                           \\
    \cohptype[n]                        & \unitpoint                              & \defeq & \unittype                                                    \\
    \cohptype[n][\lbrack 0,p \rbrack]   & (\fulldepscohs[n])                      & \defeq & \cohptype[n][\lbrack 0,p-1 \rbrack](\fulldepscohs[n]) \times \\
                                        &                                         &        & \left\{
    \begin{array}{l}
      \Pi q \leq n-p.\,\Pi r\leq q.\,\Pi \epsilon.\,\Pi \omega.                                                                                                    \\
      \Pi d:\framep[n+2][p](\overline{\deps}(\fulldepscohs.\depscohs[n][\lbrack 0,p-1 \rbrack])).\,\Pi c:\painting[n+2][p](\overline{\fulldeps}(\fulldepscohs[n])) \\
      \fulldepscohs.\depscohs.\restrp[n][p][q][\epsilon](\restrf[n+1][p](d))(\restrp[n+1][p][r][\omega]
      (\fulldepscohs[n])(d)(c)) =                                                                                                                                  \\
      \fulldepscohs.\depscohs.\restrp[n][p][r][\omega](\restrf[n+1][p](d))(\restrp[n+1][p][q+1][\epsilon]
      (\fulldepscohs[n])(d)(c))
    \end{array}
    \right.                                                                                                                                               \\
  \end{array}
\end{equation*}

\subsection{Defining frame coherence laws and specifying frame 2-dimensional coherence laws}

Thanks to $\cohptype[n][\lbrack 0,p-1 \rbrack]$, we can mutually define $\cohf[n+1][\lbrack 0,p \rbrack]$ and $\cohttype[n][\lbrack 0,p-1 \rbrack]$ for $p \leq n+1$ from $\fulldepscohs[n]$, abbreviating this and the dependency in a $\coht[n][\lbrack 0,p-1 \rbrack]$ as $\depscohttype[n][\lbrack 0,p-1 \rbrack]$, and using also abbreviations in passing for $\cohl$ and $\cohltype$. It is also convenient to build $\depscohs[n+1][\lbrack 0,p \rbrack]$ from $\depscoht[n][\lbrack 0,p-1 \rbrack]$, so that we can now build $\framep[n+3][p]$. Thus, we also include this abbreviation in the definition:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \depscohttype[n][\lbrack 0,p-1 \rbrack]                                       &                                                                               & \defeq                                                                            &
    \left(\begin{array}{lll}
              \fulldepscohs[n]                & : & \fulldepscohstype[n]                                  \\
              \cohp[n][\lbrack 0,p-1 \rbrack] & : & \cohptype[n][\lbrack 0,p-1 \rbrack](\fulldepscohs[n]) \\
            \end{array}
    \right)                                                                                                                                                                                                                                                                                     \\
    \overline{\depscohs}                                                          & (\depscoht[n][\lbrack 0,p-1 \rbrack]:\depscohttype[n][\lbrack 0,p-1 \rbrack]) & :                                                                                 & \depscohstype[n+1][\lbrack 0,p \rbrack] \\
    \overline{\depscohs}                                                          & \depscoht[n][\lbrack 0,p-1 \rbrack]                                           & \defeq                                                                            &
    \left(\begin{array}{l}
              \overline{\deps}(\depscoht.\fulldepscohs.\depscohs[n][\lbrack 0,p-1 \rbrack]) \\
              \restrp[n+1][p](\depscoht.\fulldepscohs[n])                                   \\
              \cohf[n+1][p](\depscoht[n][\lbrack 0,p-1 \rbrack])                            \\
            \end{array}\right)                                                                                                                                                                                                         \\
    \cohf[n+1][\lbrack 0,p \rbrack]                                               &
    (\depscoht[n][\lbrack 0,p-1 \rbrack]:\depscohttype[n][\lbrack 0,p-1 \rbrack]) & :                                                                             & \cohftype[n+1][\lbrack 0,p \rbrack]
    \left(\begin{array}{l}
              \overline{\fulldeps}(\depscoht.\fulldepscohs[n])              \\
              \restrp[n+1][\lbrack 0,p \rbrack](\depscoht.\fulldepscohs[n]) \\
            \end{array}\right)                                                                                                                                                                                                                         \\
    \\
    \cohf[n][\lbrack 0,0 \rbrack]                                                 &
    \unitpoint                                                                    & \defeq                                                                        & (\unitpoint,\lambda q.\lambda r. \lambda\epsilon. \lambda\omega.\lambda \_.\refl)                                           \\
    \cohf[n][\lbrack 0,p+1 \rbrack]                                               &
    (\depscoht[n][\lbrack 0,p-1 \rbrack])                                         & \defeq                                                                        &
    \left(\begin{array}{l}
              \cohf[n+1][\lbrack 0,p \rbrack]
              (\depscoht[n]),                                                    \\
              \lambda q.\lambda r.\lambda \epsilon.\lambda \omega.\lambda (d,l). \\
              (\cohf[n][\lbrack 0,p \rbrack][q][r][\epsilon][\omega]
              (\depscoht[n])(d), \cohl[n][p][q][r][\epsilon][\omega]
              (\depscoht[n])(d)(l))                                              \\
            \end{array}\right)                                                                                                                                                                                                                    \\
    \coht[n][p]                                                                   &
    \fulldepscohs[n]                                                              & :                                                                             & \cohttype[n][p](\fulldepscohs[n])                                                                                           \\
    \coht[n][p][q][r][\epsilon][\omega][\theta]                                   &
    \fulldepscohs[n]                                                              & \defeq                                                                        & \UIP                                                                                                                        \\

    \cohl[n][p]                                                                   &
    (\depscoht[n][\lbrack 0,p-1 \rbrack]:\depscohttype[n][\lbrack 0,p-1 \rbrack]) & :                                                                             &
    \begin{array}{l}
      \Pi q \leq n-p.\,\Pi r\leq q.\,\Pi \epsilon.\,\Pi \omega.                                                                          \\
      \Pi d.\Pi l. \overrightarrow{\cohf[n][p][r+1][q+1][\epsilon][\omega]}{(\restrl[n][p][q][\epsilon](\restrl[n][p][r][\omega](l)))} = \\
      \restrl[n][p][r][\omega](\restrl[n][p][q+1][\epsilon](l))
    \end{array}                                                                                                                                                          \\
    \cohl[n][p][q][r][\epsilon][\omega]                                           &
    (\depscoht[n][\lbrack 0,p-1 \rbrack])                                         & \defeq                                                                        &
    \lambda d. \lambda l. \lambda \theta. \overrightarrow{\coht[n][p][q][r][\epsilon][\omega][\theta](\depscoht.\fulldepscohs[n])}{(\depscoht.\cohp[n][p][q][r][\epsilon][\omega](l_\theta))}                                                                                                   \\
  \end{array}
\end{equation*}
where $\cohttype$ is:
\begin{equation*}
  \cohttype[n][p](\fulldepscohs[n]) \defeq
  \begin{array}{l}
    \Pi q \leq n-p.\,\Pi r \leq q.\,\Pi \epsilon.\,\Pi \omega.\,\Pi \theta.                         \\
    \fulldepscohs.\depscohs.\cohf[n][p][r][0][\omega][\theta](\restrf[n+1][p][q+2][\epsilon]) \circ \\
    \ap\restrf[n][p][r][\omega](\cohf[n+1][p][q+1][0][\epsilon][\theta]
    (\fulldepscoht[n]))
    \circ                                                                                           \\
    \fulldepscohs.\depscohs.\cohf[n][p][q][r][\epsilon][\omega](\restrf[n+1][p][0][\theta]) =       \\
    \ap\restrf[n][p][0][\theta](\cohf[n+1][p][q+1][r+1][\epsilon][\omega]
    (\fulldepscoht[n]))
    \circ                                                                                           \\
    \fulldepscohs.\depscohs.\cohf[n][p][q][0][\epsilon][\theta](\restrf[n+1][p][r+1][\omega])
    \circ                                                                                           \\
    \ap\restrf[n][p][q][\epsilon](\cohf[n+1][p][r][0][\omega][\theta]
    (\fulldepscoht[n]))
  \end{array}
\end{equation*}

It is also convenient to introduce a notion of full dependencies at the level of the coherence conditions of dimension 2, that combines a full $\depscoht[n][\lbrack 0,n \rbrack]$ with a set $E:\framep[n+3][n+3](\overline{\deps}(\overline{\depscohs}(\depscoht[n][\lbrack 0,n \rbrack]))) \rightarrow \U$:
\begin{equation*}
  \fulldepscohttype[n] \defeq
  \left(\begin{array}{lll}
      \depscoht[n][\lbrack 0,n \rbrack] & : & \depscohttype[n][\lbrack 0,n \rbrack]                                                                      \\
      E                                 & : & \framep[n+3][n+3](\overline{\deps}(\overline{\depscohs}(\depscoht[n][\lbrack 0,n \rbrack])))\rightarrow \U \\
    \end{array}\right)
\end{equation*}

We can then extend $\overline{\depscohs}$ to build $\fulldepscohs$ from $\fulldepscoht$:

\begin{align*}
  \overline{\fulldepscohs}(\fulldepscoht[n]:\fulldepscohstype[n]) & :     \fulldepscohstype[n+1]                                                       \\
  \overline{\fulldepscohs}(\fulldepscoht[n])                      & \defeq \left(\begin{array}{l}
                                                                                     \overline{\depscohs}(\fulldepscoht.\depscoht[n][\lbrack 0,n \rbrack]) \\
                                                                                     \fulldepscoht.E                                                       \\
                                                                                   \end{array}\right) \\
\end{align*}

\subsection{Defining painting coherence laws}

The last key construction we need is $\cohp[n+1][p]$ for $p \leq n+1$, which is by recursion on $n-p$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \cohp[n+1][p]                                &
    (\fulldepscoht[n]:\fulldepscohttype[n])      & :      &
    \cohptype[n+1][p](\overline{\fulldepscohs}(\fulldepscoht[n]))                                       \\
    \cohp[n+1][n+1][q][r][\epsilon][\omega]      &
    (\fulldepscoht[n])
                                                 & \defeq & \lambda d. \lambda (l,c). l\epsilon         \\
    \cohp[n+1][p \leq n][q][r][\epsilon][\omega] &
    (\fulldepscoht[n])
                                                 & \defeq & \lambda d. \lambda (l,c).
    \left(\begin{array}{l}
              \cohl[n][p][q][r][\epsilon][\omega](\fulldepscoht.\depscoht[n][\lbrack 0,p-1 \rbrack])(d)(l), \\
              \cohp[n+1][p+1][q][r][\epsilon][\omega](\fulldepscoht[n])(d,l))(c)                            \\
            \end{array}\right) \\
  \end{array}
\end{equation*}

From which we can deduce $\cohp[n+1][\lbrack 0,p-1 \rbrack]$ for $p \leq n+2$ again by a second recursion, this time on $p$:

\begin{equation*}
  \begin{array}{llcl}
    \cohp[n+1][\lbrack 0,p-1 \rbrack]       &
    (\fulldepscoht[n]:\fulldepscohttype[n]) & :      & \cohptype[n+1][\lbrack 0,p-1 \rbrack](\fulldepscoht.\depscoht.\fulldepscohs[n])) \\
    \cohp[n+1]                              &
    \fulldepscohs[n]                        & \defeq & \unittype                                                                        \\
    \cohp[n+1][\lbrack 0,p \rbrack]         &
    \fulldepscoht[n]
                                            & \defeq &
    (\cohp[n+1][\lbrack 0,p-1 \rbrack](\fulldepscoht[n]),\cohp[n+1][p](\fulldepscoht[n]))                                               \\
  \end{array}
\end{equation*}

\subsection{Completing the construction by induction}

Armed with all the constructions, we are now able to specify the whole
structure on which we can reason by induction, for $D:\Xto{m}{n}$
given:

\begin{align*}
  \framep[n][\lbrack 0,n \rbrack]                                                                      & :      & \frametype[n][\lbrack 0,n \rbrack]                                     \\
  \painting[n][\lbrack 0,n \rbrack]                                                                    & :      & \paintingtype[n][\lbrack 0,n \rbrack](\framep[n][\lbrack 0,n \rbrack]) \\
  \restrf[n][\lbrack 0,n \rbrack]                                                                      & :      & \restrftype[n][\lbrack 0,n \rbrack]
  \left(
  \begin{array}{l}
      \framep[n][\lbrack 0,n \rbrack]   \\
      \painting[n][\lbrack 0,n \rbrack] \\
    \end{array}\right)                                                                                                                                                      \\
  \deps[n][\lbrack 0,n \rbrack]                                                                        & \defeq &
  \left(
  \begin{array}{l}
      \framep[n][\lbrack 0,n \rbrack]   \\
      \painting[n][\lbrack 0,n \rbrack] \\
      \restrf[n][\lbrack 0,n \rbrack]   \\
    \end{array}\right)                                                                                                                                                      \\
  \fulldeps[n](E: \framep[n+1][n+1](\deps[n][\lbrack 0,n \rbrack])\rightarrow \U)                      & \defeq &
  \left(
  \begin{array}{l}
      \deps[n][\lbrack 0,n \rbrack] \\
      E                             \\
    \end{array}\right)                                                                                                                                                          \\
  \restrp[n][\lbrack 0,n \rbrack](E)                                                                   & :      & \restrptype[n][\lbrack 0,n \rbrack]
  (\fulldeps[n](E))                                                                                                                                                                      \\
  \cohf[n][\lbrack 0,n \rbrack](E)                                                                     & :      & \cohftype[n][\lbrack 0,n \rbrack]
  \left(
  \begin{array}{l}
      \fulldeps[n](E)                    \\
      \restrp[n][\lbrack 0,n \rbrack](E) \\
    \end{array}\right)                                                                                                                                                     \\
  \depscohs[n][\lbrack 0,n \rbrack](E: \framep[n+1][n+1](\deps[n][\lbrack 0,n \rbrack])\rightarrow \U) & \defeq &
  \left(
  \begin{array}{l}
      \fulldeps[n](E)                    \\
      \restrp[n][\lbrack 0,n \rbrack](E) \\
      \cohf[n][\lbrack 0,n \rbrack](E)   \\
    \end{array}\right)                                                                                                                                                     \\
  \fulldepscohs[n]
  \left(\begin{array}{l}
            E: \framep[n+1][n+1](\deps[n][\lbrack 0,n \rbrack])\rightarrow \U                           \\
            E': \framep[n+2][n+2](\overline{\deps}(\depscohs[n][\lbrack 0,n \rbrack](E)))\rightarrow \U \\
          \end{array}\right)       & \defeq &
  \left(
  \begin{array}{l}
      \depscohs[n][\lbrack 0,n \rbrack](E) \\
      E'                                   \\
    \end{array}\right)                                                                                                                                                   \\
  \cohp[n][\lbrack 0,n \rbrack](E)(E')                                                                 & :      & \cohptype[n][\lbrack 0,n \rbrack]
  (\fulldepscohs[n](E)(E'))                                                                                                                                                              \\
\end{align*}

Indeed, to build this structure at level $n+1$, from level $n$, it is
enough to assume $E:\framep[n+1][n+1](\deps[n][\lbrack 0,n \rbrack]) \rightarrow \U$ (where $\deps[n][\lbrack 0,n \rbrack]$ come from the assumptions at level $n$) and to build the following:
\begin{itemize}
  \item $\framep[n+1][\lbrack 0,n+1 \rbrack]$ comes from $\deps[n][\lbrack 0,n \rbrack]$.
  \item $\painting[n+1][\lbrack 0,n+1 \rbrack]$ comes from $\fulldeps[n](E)$.
  \item $\restrf[n+1][\lbrack 0,n+1 \rbrack]$ comes from $\depscohs[n][\lbrack 0,n \rbrack](E)$.
  \item $\restrp[n+1][\lbrack 0,n+1 \rbrack](E')$, for $E':\framep[n+2][n+2](\overline{\deps}(\depscohs[n][\lbrack 0,n \rbrack])) \rightarrow \U$, comes from $\fulldepscohs[n](E)(E')$.
  \item $\depscoht[n][\lbrack 0,n \rbrack](E)(E')$ comes from $\fulldepscohs[n](E)(E')$ and $\cohp[n][\lbrack 0,n \rbrack](E)(E')$.
  \item $\cohf[n+1][\lbrack 0,n+1 \rbrack](E')$ comes from $\depscoht[n][\lbrack 0,n \rbrack](E)(E')$.
  \item $\fulldepscoht[n](E)(E')(E'')$, for $E'': \framep[n+3][n+3](\overline{\deps}(\overline{\depscohs}(\depscoht[n][\lbrack 0,n \rbrack]))) \rightarrow \U$, comes from $\depscoht[n][\lbrack 0,n \rbrack](E)(E')$ and $E''$.
  \item $\cohp[n+1][\lbrack 0,n+1 \rbrack](E')(E'')$ comes from $\fulldepscoht[n](E)(E')(E'')$.
\end{itemize}

Therefore, assuming a construction of $\framep[n][n]$ from $D:\Xto{m}{n}$, we are able to build a $\framep[n+1][n+1]$ from $D$ extended with some $E:\framep[n+1][n+1](\deps[n][\lbrack 0,n \rbrack]) \rightarrow \U$. This is enough to build $\Xcomp{m}{n}(D)$, for $D:\Xto{m}{n}$, and thus $\Xp{m}$, by starting with an empty $D$.

\bibliographystyle{msclike}
\bibliography{paper}

\end{document}